### 2014-12-10 paper.tex
Initial final report checkin

\documentclass{article}

\usepackage{fullpage}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{bold-extra}

\titleformat{\section}{\normalsize\bf}{}{0em}{}

\begin{document}

%\rightline{ }
\centerline{\large{\textbf{Distributed Multi-Heuristic A* (HAMSTAR)}}}
\centerline{Noam Brown, Aram Ebtekar, Yuzuko Nakamura}
\centerline{15-712 Fall 2014 Final Project Report}

\section{Abstract}

A* is a popular graph search algorithm used in many artificial intelligence applications. The original A* algorithm makes use of a single heuristic to prioritize exploration through a graph, but a variation on the algorithm known as multi-heuristic A* (MHA*) makes use of multiple heuristics to more quickly arrive at approximate solutions. For this project we implemented a distributed version of MHA*, with one process per heuristic. We show... [performance]
%TODO: Highlight important results

\section{Introduction}

Weighted A* is a simple, heuristic-based search algorithm used in artificial intelligence applications such as robotics and games. A heuristic function estimates the remaining distance from any state to the goal, thus guiding the search toward the goal. If the heuristic is \textbf{admissible}, meaning it never overestimates the true distance, then the solution is guaranteed to be optimal to within the weight factor. If it meets a stronger condition called \textbf{admissibility}, then no state needs to be expanded more than once.

Finding good heuristic functions is difficult, and moreover, finding an admissible heuristic function that is reasonably accurate over the entirety of the search space is often not practical. Multi-heuristic A* (MHA*) \cite{Aine14} is an alteration of the A* algorithm that allows the usage of a combination of multiple arbitrary heuristics to guide the search, while retaining the optimality bound of wA* provided that one consistent heuristic is used as an “anchor”. Where the shortcomings of each individual heuristic may result in local optima which trap the search, the ability to change heuristics can provide an escape.

In their work, Aine et al. propose two variants on MHA*: Independent Multi-Heuristic A* (IMHA*), where path cost and shortest path information are tracked separately for each heuristic, and Shared Multi-Heuristic A* (SMHA*), where shared knowledge of path cost and shortest path information are updated by all heuristics. SMHA* is the more effective of the two (explores the graph in a less repetitive manner and enables more co-operation between heuristics), but is more difficult to parallelize due to the amount of shared data.

For this project, we implemented a distributed version of this second variant, SMHA*.

\section{Relevant Work}

Our work is directly based off of Aine et al.'s original multi-heuristic anchored search \cite{Aine14} mentioned above. A parallelized version of the original A* algorithm was presented by Phillips et al. in \cite{Phillips14}. This method of parallelization makes inferences on when it's safe to expand multiple nodes in parallel, and in so doing maintains the approximate suboptimality guarantees of sequential wA*. This parallelization is dependent on the expansion operation being expensive enough to justify expanding multiple nodes in parallel, which may not be the case for all problems. By contrast, our method parallelizes the search process by constantly expanding multiple nodes chosen by different heuristics (instead of multiple nodes chosen by a single heuristic). As a result, our method differs from the previous in that, although nodes may be expanded multiple times by different machines, we do not require expansions to be slow.

\section{System Description}

\begin{figure}
\centering \includegraphics[width=5.0in]{system-diagram}
\caption{A diagram of our system. One machine starts up the others and runs the admissible anchor search. The other machines run their individual heuristics. Communication is done using MPI. 
\textcircled{1}\textbf{Anchor:} Starts heuristic machines; performs state expansions using anchor heuristic; receives, resolves, and broadcasts table updates; checks for termination conditions. \textcircled{2}\textbf{Heuristic:} Expands likely state (chosen using heuristic); updates path cost to its successors; communicate updates to anchor; termination conditions}

\end{figure}

\section{Experimental Setup}

%\section{Goals}

%\subsection{75\% Goal}
%\begin{enumerate}
%\item Implement parallel versions of IMHA* and SMHA*
%\item Identify bottlenecks in performance in running SMHA*
%\end{enumerate}

%\subsection{100\% Goal}
%\begin{enumerate}
%\item Implement distributed versions of IMHA* and SMHA*, perhaps with fancier distributed heuristic selection (there exists recent work which uses a simple form of online learning)
%\end{enumerate}

%\subsection{125\% Goal}
%\begin{enumerate}
%\item Design and implement a more sophisticated intermediate between IMHA* and SMHA*, based on the analysis of parallel and distributed system tradeoffs from our preliminary evaluations.
%\end{enumerate}

%\section{Resources Needed}

%We will probably need a machine with at least 32 cores for the 75\% goal and a cluster of some number of machines for the distributed version (100\% goal). We will be using MPI for message-passing between computers and C++ to program.

%\section{Relevance to the Class}

%We will investigate ways to parallelize SMHA*, and compare its performance to parallelized IMHA*. This project gives us the opportunity to explore classic tradeoffs in designing a distributed system, particularly between network latency and local computation.

%\section{Work Distribution}

%Tasks for 75\% goal:
%\begin{enumerate}
%\item Implement IMHA* using locks / shared memory
%\item Implement SMHA* using locks / shared memory
%\item Pull problem and heuristics from cited work
%\item Run algorithms on at least one of the cited example problems
%\item Evaluate performance overhead of the parallelism
%\end{enumerate}

%Tasks for 100\% goal:
%\begin{enumerate}
%\item Implement IMHA*/SHMA* using message passing
%\item Run algorithms on a cluster of machines
%\item Evaluate performance overhead of local and networked message passing
%\item Reflect and compare against multicore version
%\end{enumerate}

%Tasks for 125\% goal:
%\begin{enumerate}
%\item Implement new version of SMHA* based on considerations of the evaluation
%\item Test and evaluate
%\end{enumerate}
%\newpage

%\section{Final Paper Outline}

%\subsection{Introduction}

%Similar to the above problem statement.

%\subsection{Related Work}

%\cite{Aine14} (MHA*) presents the original multi-heuristic anchored search. \cite{Phillips14} (parallelized A* for slow expansions) presents a criterion for parallel expansion using locks, while retaining the approximate suboptimality guarantees of sequential wA*.

%\subsection{System Description}

%A distributed system running one heuristic on each machine.

%\subsection{Evaluation}

%\textbf{IMHA* experiment:}

%We will implement IMHA* on 32 cores, each running its own heuristic.

%\textbf{SMHA* experiment:}

%We will implement SMHA* on 32 cores, each running its own heuristic and updating shared data structures for path costs.

%We want to keep the approximate optimality guarantees of unparallelized MHA*, so this requires careful maintenance of the data structures to ensure the invariants are kept.

%Explore different way of updating the shared data (i.e. data replication and consistency checking/message-passing?)?

%Performance of the various implementations will be evaluated on the problems from Aine et al.:
%\begin{enumerate}
%\item Robot arm manipulation
%\item 3D path planning (navigation)
%\item Sliding tile puzzle
%\end{enumerate}

\subsection{Conclusions and Future Work}

\bibliographystyle{ieeetr}
\bibliography{sources}

\end{document}

### 2014-12-10 SMHA_time.cpp
Added stats stuff

@@ -32,7 +32,7 @@ constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
 
 //communication constants
-constexpr bool COMM_ITER = false;
+constexpr bool COMM_ITER = true;
 constexpr double COMM_INTERVAL = 0.05;
 constexpr int COMM_FREQ = 1000;
 constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*4; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
@@ -476,7 +476,7 @@ public:
           int index = 0;
           memcpy(&child_buffer[index++], &opt_bound, sizeof(Cost));
           child_buffer[index++] = updated_states.size();
-		  assert(updated_states.size() <= BUFFER_SIZE);
+		  assert(updated_states.size() <= BUFFER_SIZE*comm_size);
 		  child_buffer[index++] = total_discovered;
 		  child_buffer[index++] = total_expanded;
 		  child_buffer[index++] = finished;

### 2014-12-10 SMHA.cpp
Frequency experiments

@@ -32,9 +32,9 @@ constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
 
 //communication constants
-constexpr bool COMM_ITER = false;
+constexpr bool COMM_ITER = true;
 constexpr double COMM_INTERVAL = 0.05;
-constexpr int COMM_FREQ = 1000;
+constexpr int COMM_FREQ = 1;
 constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*4; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
 constexpr int DATUM_SIZE = (GRID_ROWS * GRID_COLS)*2 + 1 + 2; //4x4 state, backtrace, mask, g, and h. Assumes Cost is same size as int
 constexpr int HEAD_NODE = 0;
@@ -215,7 +215,6 @@ public:
   // id of the Searcher
   int id;
   // simulate network communication with message queue and corresponding mutexes
-  vector<queue<State> > network;
 
   //Communication stuff
   int * head_buffer;

### 2014-12-10 SMHA_time.cpp

@@ -32,7 +32,7 @@ constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
 
 //communication constants
-constexpr bool COMM_ITER = true;
+constexpr bool COMM_ITER = false;
 constexpr double COMM_INTERVAL = 0.05;
 constexpr int COMM_FREQ = 1000;
 constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*4; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.

### 2014-12-10 SMHA.cpp
More frequencies

@@ -34,7 +34,7 @@ constexpr Cost INFINITE = 1e30;
 //communication constants
 constexpr bool COMM_ITER = true;
 constexpr double COMM_INTERVAL = 0.05;
-constexpr int COMM_FREQ = 1;
+constexpr int COMM_FREQ = 64000;
 constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*4; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
 constexpr int DATUM_SIZE = (GRID_ROWS * GRID_COLS)*2 + 1 + 2; //4x4 state, backtrace, mask, g, and h. Assumes Cost is same size as int
 constexpr int HEAD_NODE = 0;

### 2014-12-10 SMHA.cpp
Correct frequencies

@@ -34,7 +34,7 @@ constexpr Cost INFINITE = 1e30;
 //communication constants
 constexpr bool COMM_ITER = true;
 constexpr double COMM_INTERVAL = 0.05;
-constexpr int COMM_FREQ = 64000;
+constexpr int COMM_FREQ = 1000;
 constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*4; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
 constexpr int DATUM_SIZE = (GRID_ROWS * GRID_COLS)*2 + 1 + 2; //4x4 state, backtrace, mask, g, and h. Assumes Cost is same size as int
 constexpr int HEAD_NODE = 0;
@@ -417,7 +417,7 @@ public:
           opt_bound = max(opt_bound, open.cbegin()->first);
         }
       }
-	  if ((COMM_ITER && iter % COMM_FREQ == 0) || (!COMM_ITER && iter % 10 == 0 && chrono::duration<double, chrono::seconds::period>(Clock::now() - last_time).count() > COMM_INTERVAL)) {
+	  if ((COMM_ITER && iter % 50 == 0) || (!COMM_ITER && iter % 10 == 0 && chrono::duration<double, chrono::seconds::period>(Clock::now() - last_time).count() > COMM_INTERVAL)) {
 		  last_time = Clock::now();
         //cout << "Process " << comm_rank << " preparing to communicate " << endl;
         //Handle communication

### 2014-12-10 SMHA.cpp
began lame_SMHA.cpp

@@ -212,9 +212,6 @@ public:
   int num_expanded;
   int total_discovered;
   int total_expanded;
-  // id of the Searcher
-  int id;
-  // simulate network communication with message queue and corresponding mutexes
 
   //Communication stuff
   int * head_buffer;
@@ -280,8 +277,8 @@ public:
     // the start state is discovered but not yet expanded
     num_discovered = 1;
     num_expanded = 0;
-	total_discovered = 0;
-	total_expanded = 0;
+    total_discovered = 1;
+    total_expanded = 0;
     start_time = Clock::now();
   }
 
@@ -401,7 +398,7 @@ public:
 
     cout << "Starting run" << endl;
     int iter = 0;
-	Clock::time_point last_time = Clock::now();
+    Clock::time_point last_time = Clock::now();
     // repeat until some thread declares the search to be finished
     while (true) {
 		if (num_expanded > benchmark) {

### 2014-12-10 SMHA.cpp
c++ is broken

@@ -225,6 +225,7 @@ public:
     Cost hMT = misplacedTiles(s, goal);
     s_data.hAnch = hMD + hLC;
     s_data.h = MD*hMD + LC*hLC + MT*hMT;
+    s_data.iter = open.cend();
   }
 
   // key for the priority queue
@@ -239,8 +240,10 @@ public:
   }
 
   void erase(State& s, StateData& s_data) {
-    open.erase(s_data.iter);
-    s_data.iter = open.cend();
+    if (s_data.iter != open.cend()) {
+      open.erase(s_data.iter);
+      s_data.iter = open.cend();
+    }
   }
 
   // assumes start, goal, w1 and w2 are already set
@@ -269,7 +272,6 @@ public:
     StateData& start_data = data[start];
     start_data.g = 0;
     computeH(start, start_data);
-    start_data.iter = open.cend();
     start_data.bp = start.first;
     insert(start, start_data);
 
@@ -290,7 +292,6 @@ public:
       StateData& t_data = data[t];
       if (t_data.g == INFINITE) {
         computeH(t, t_data);
-        t_data.iter = open.cend();
         num_discovered++;
 		total_discovered++;
       }
@@ -347,11 +348,13 @@ public:
     s_data.mask |= maskNew;
     // if necessary, compute heuristics
     if (s_data.g == INFINITE) {
-      s_data.iter = open.cend();
-      if (comm_rank == HEAD_NODE)
+      if (comm_rank == HEAD_NODE) {
         s_data.h = s_data.hAnch = hAnchNew;
-      else
+        s_data.iter = open.cend();
+      }
+      else {
         computeH(s, s_data);
+      }
     }
     // if necessary, close this state
     if ((s_data.mask & closing_mask) && s_data.iter != open.cend()) {
@@ -512,8 +515,12 @@ public:
       // get a State to expand
       auto it = open.cbegin();
 
+      // timing
+      Clock::time_point cur_time = Clock::now();
+      time_elapsed = chrono::duration<double, chrono::seconds::period>(cur_time - start_time).count();
+
       // search termination condition
-      if (it == open.cend() || data[goal].g <= w2 * opt_bound || time_elapsed > TIME_LIMIT) {
+      if (data[goal].g <= w2 * opt_bound) {
         // flush and kill, but don't kill the master!!!!!
 		  finished = 1;
 		  cout << "Machine " << comm_rank << " found a solution" << endl;
@@ -600,6 +607,12 @@ public:
[…]

### 2014-12-10 SMHA.cpp
fixed using header

@@ -231,7 +231,7 @@ public:
   // key for the priority queue
   Cost f(StateData& s_data) { return s_data.g + w1*s_data.h; }
 
-  void insert(State& s, StateData& s_data) {
+  void insert(const State& s, StateData& s_data) {
     // if state is already in open[id], remove it before re-inserting
     if (s_data.iter != open.cend())
       open.erase(s_data.iter);
@@ -239,7 +239,7 @@ public:
     s_data.iter = open.insert(pair<Cost, State>(f(s_data), s));
   }
 
-  void erase(State& s, StateData& s_data) {
+  void erase(const State& s, StateData& s_data) {
     if (s_data.iter != open.cend()) {
       open.erase(s_data.iter);
       s_data.iter = open.cend();

### 2014-12-10 lame_SMHA.cpp

@@ -1,0 +1,415 @@
+#include <iostream>
+#include <array>
+#include <cmath>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+#include <cassert>
+#include <vector>
+#include <queue>
+#include <set>
+#include <map>
+#include <algorithm>
+//#include <unordered_map>
+#include <cstdio>
+#include <chrono>
+#include <string>
+
+using namespace std;
+
+typedef float Cost;
+typedef chrono::high_resolution_clock Clock;
+// for a state s, s.first is the board and s.second is its inverse
+// permutation: s.first[s.second[i]] = s.second[s.first[i]] = i
+typedef pair<vector<int>, vector<int> > State;
+
+// compile-time constants
+constexpr int NUM_HEURISTICS = 8;
+constexpr int MASK_CLOSED = 1;
+constexpr int MASK_CLOSED_ANCHOR = 2;
+constexpr int GRID_ROWS = 5;
+constexpr int GRID_COLS = 5;
+constexpr int NUM_MOVES = 4;
+constexpr double TIME_LIMIT = 300;
+constexpr Cost INFINITE = 1e30;
+constexpr int HEAD_NODE = 0;
+
+#include "lame_SMHA.h"
+
+static int finished = -1;
+int SEED = 1;
+string filename = "test";
+
+// right, up, left, down
+constexpr array<int, NUM_MOVES> moves = { 1, -GRID_COLS, -1, GRID_COLS };
+
+// print a nicely formatted board
+void printState(const State& s) {
+  for (int i = 0; i < GRID_ROWS; ++i) {
+    for (int j = 0; j < GRID_COLS; ++j)
+      printf("%3d", s.first[i*GRID_COLS + j]);
+    printf("\n");
+  }
+  printf("\n");
+}
+
+Cost manhattanDist(const State& s1, const State& s2) {
+  Cost h = 0;
+  for (int val = 1; val < GRID_ROWS*GRID_COLS; ++val) {
+    // compute distance between val's position in s1 and s2
+    int r1 = s1.second[val] / GRID_COLS;
+    int c1 = s1.second[val] % GRID_COLS;
+    int r2 = s2.second[val] / GRID_COLS;
+    int c2 = s2.second[val] % GRID_COLS;
+    h += abs(r1 - r2) + abs(c1 - c2);
+  }
+  return h;
+}
+
+Cost misplacedTiles(const State& s1, const State& s2) {
[…]

### 2014-12-10 lame_SMHA.h

@@ -1,0 +1,51 @@
+
+// data associated with one state
+class StateData {
+public:
+  // back-pointer to previous state along the discovered path
+  vector<int> bp;
+  // stores Boolean flags such as whether the state is CLOSED
+  int mask;
+  // g is the cost of a discovered path, h estimates the remaining cost to goal
+  Cost g;
+  Cost h[NUM_HEURISTICS];
+  // points to the state's location in the priority queue
+  multimap<Cost, State>::const_iterator iter[NUM_HEURISTICS];
+  // initialization
+  StateData() : mask(0), g(INFINITE) {}
+};
+
+
+
+// a process which performs a search, typically assigned to one thread on one machine
+class Searcher {
+public:
+  // multi-heuristic mixing coefficients
+  double MD, LC, MT;
+  // Searcher-specific closing mask
+  int closing_mask;
+  // number of states seen and expanded
+  int num_discovered;
+  int num_expanded;
+  // id of the Searcher
+  int comm_rank;
+  // the priority queue or search frontier
+  multimap<Cost, State> open;
+
+  // key for the priority queue
+  Cost f(StateData& s_data);
+
+  void insert(const State& s, StateData& s_data);
+
+  void erase(const State& s, StateData& s_data);
+
+  // assumes start, goal, w1 and w2 are already set
+  void init();
+
+  void expand(const State& s, StateData& s_data);
+
+  void runSingleIteration();
+
+  // get the path: a sequence of states from start to goal
+  vector<State> getSolution();
+};

### 2014-12-10 SMHA.cpp
lame works

@@ -288,12 +288,14 @@ public:
     assert(!(s_data.mask & closing_mask));
     s_data.mask |= MASK_CLOSED | closing_mask;
 
+    erase(s, s_data);
+
     for (State& t : getSuccessors(s)) {
       StateData& t_data = data[t];
       if (t_data.g == INFINITE) {
         computeH(t, t_data);
         num_discovered++;
-		total_discovered++;
+        total_discovered++;
       }
 
       // critical section for updating g-value and inserting
@@ -617,9 +619,8 @@ public:
       }
       State s = it->second;
       StateData& s_data = data[s];
-      erase(s, s_data);
       num_expanded++;
-	  total_expanded++;
+      total_expanded++;
       expand(s, s_data);
 
       iter++;
@@ -713,12 +714,12 @@ int main(int argc, char** argv) {
 	if (comm_rank == HEAD_NODE) {
 		Cost path_length = searcher.data[searcher.goal].g;
 		// print solution if it was found
-		/*if (path_length <= searcher.w2 * searcher.opt_bound) {
-			for (State& s : searcher.getSolution()) {
+		if (path_length <= searcher.w2 * searcher.opt_bound) {
+			/*for (State& s : searcher.getSolution()) {
 				printState(s);
-			}
-		}*/
-		cout << "Found path of length " << path_length << " Discovered nodes = " << searcher.num_discovered << ". Expanded nodes: " << searcher.num_expanded << ". Time: " << searcher.time_elapsed << endl;
+			}*/
+      cout << "Found path of length " << path_length << " Discovered nodes = " << searcher.num_discovered << ". Expanded nodes: " << searcher.num_expanded << ". Time: " << searcher.time_elapsed << endl;
+		}
 		cout << "Total discovered: " << searcher.total_discovered << " total expanded: " << searcher.total_expanded << endl;
 		fprintf(fout, "%f %f %f %d %f\n", searcher.w1, searcher.w2, searcher.time_elapsed, searcher.num_expanded, path_length);
 	}

### 2014-12-10 lame_SMHA.cpp

@@ -265,6 +265,14 @@ void Searcher::expand(const State& s, StateData& s_data) {
 }
 
 void Searcher::runSingleIteration() {
+  if (comm_rank == HEAD_NODE) {
+    if (open.empty()) {
+      opt_bound = INFINITE;
+    }
+    else {
+      opt_bound = max(opt_bound, open.cbegin()->first);
+    }
+  }
   // get a State to expand
   auto it = open.cbegin();
 
@@ -279,7 +287,6 @@ void Searcher::runSingleIteration() {
 
   State s = it->second;
   StateData& s_data = data[s];
-  erase(s, s_data);
   num_expanded++;
   total_expanded++;
   expand(s, s_data);
@@ -333,11 +340,6 @@ void prepareDistributedSearch() {
 
   // initialize global data structures
   data.clear();
-  // create data entry for start state
-  StateData& start_data = data[start];
-  start_data.g = 0;
-  computeH(start, start_data);
-  start_data.bp = start.first;
   opt_bound = 0;
   // the start state is discovered but not yet expanded
   total_discovered = 1;
@@ -345,6 +347,12 @@ void prepareDistributedSearch() {
 
   vec_search.clear();
   vec_search.resize(NUM_HEURISTICS);
+  // create data entry for start state
+  StateData& start_data = data[start];
+  start_data.g = 0;
+  computeH(start, start_data);
+  start_data.bp = start.first;
+
   for (int i = 0; i < NUM_HEURISTICS; ++i) {
     vec_search[i].comm_rank = i;
     vec_search[i].open.clear();
@@ -372,6 +380,10 @@ void runDistributedSearch() {
     Clock::time_point cur_time = Clock::now();
     time_elapsed = chrono::duration<double,chrono::seconds::period>(cur_time-start_time).count();
     iter++;
+    if (total_expanded > benchmark) {
+      benchmark += 100000;
+      cout << "Total expanded " << total_expanded << endl;
+    }
   }
 }
 
@@ -402,8 +414,8 @@ int main(int argc, char** argv) {
 			for (State& s : vec_search[finished].getSolution()) {
 				printState(s);
 			}
+      cout << "Found path of length " << path_length << /*" Discovered nodes = " << searcher.num_discovered << ". Expanded nodes: " << searcher.num_expanded << ". Time: " << */time_elapsed << endl;
 		}
[…]

### 2014-12-10 lame_SMHA.cpp
Lame 16 / 4 heuristics results

@@ -24,7 +24,7 @@ typedef chrono::high_resolution_clock Clock;
 typedef pair<vector<int>, vector<int> > State;
 
 // compile-time constants
-constexpr int NUM_HEURISTICS = 8;
+constexpr int NUM_HEURISTICS = 16;
 constexpr int MASK_CLOSED = 1;
 constexpr int MASK_CLOSED_ANCHOR = 2;
 constexpr int GRID_ROWS = 5;

### 2014-12-10 lame_SMHA.cpp
mystery bug: this->erase crashes while vec_search[comm_rank].erase runs fine

@@ -227,7 +227,6 @@ void Searcher::init() {
     MT = dis(gen2);
     closing_mask = MASK_CLOSED;
   }
-
   num_discovered = 1;
   num_expanded = 0;
 }
@@ -411,10 +410,10 @@ int main(int argc, char** argv) {
 		Cost path_length = data[goal].g;
 		// print solution if it was found
 		if (path_length <= w2 * opt_bound) {
-			for (State& s : vec_search[finished].getSolution()) {
+			/*for (State& s : vec_search[finished].getSolution()) {
 				printState(s);
-			}
-      cout << "Found path of length " << path_length << /*" Discovered nodes = " << searcher.num_discovered << ". Expanded nodes: " << searcher.num_expanded << ". Time: " << */time_elapsed << endl;
+			}*/
+      cout << "Found path of length " << path_length << /*" Discovered nodes = " << searcher.num_discovered << ". Expanded nodes: " << searcher.num_expanded << */". Time: " << time_elapsed << endl;
 		}
 		cout << "Total discovered: " << total_discovered << " total expanded: " << total_expanded << endl;
 

### 2014-12-10 paper.tex
Fleshing out paper, adding the graphs I generated

@@
 
-\titleformat{\section}{\normalsize\bf}{}{0em}{}
+\titleformat{\section}{\normalsize\bf}{\thesection \hspace{5pt}}{0em}{}
+\titleformat{\subsection}{\normalsize\bf}{\thesubsection \hspace{5pt}}{0em}{}
 
@@
 
-%\rightline{ }
 \centerline{\large{\textbf{Distributed Multi-Heuristic A* (HAMSTAR)}}}
@@
 A parallelized version of the original A* algorithm was presented by Phillips et al.
-in \cite{Phillips14}.
+\cite{phillips2014pa}.
 This method of parallelization makes inferences on when it's safe to expand multiple nodes in parallel, and in so doing maintains the approximate suboptimality guarantees of sequential wA*.
@@
 \textcircled{1}\textbf{Anchor:} Starts heuristic machines; performs state expansions using anchor heuristic; receives, resolves, and broadcasts table updates; checks for termination conditions.
-\textcircled{2}\textbf{Heuristic:} Expands likely state (chosen using heuristic); updates path cost to its successors; communicate updates to anchor; termination conditions}
+\textcircled{2}\textbf{Heuristic:} Expands a likely state (chosen using heuristic); updates path cost to its successors; communicates updates to anchor; checks for termination conditions}
+\label{fig:sysdiag}
+\end{figure}
 
+An overview of our system is in Figure \ref{fig:sysdiag}.
+The main (anchor) machine starts up other heuristic machines.
+Each machine runs its own heuristic, with the anchor machine's being the admissible heuristic.
+Each of the non-anchor heuristic machines periodically send updates of the data structures (tables containing shortest path information found so far) to the anchor machine, which compares the data it receives and sends out table updates of the best paths known so far to all other nodes.
+
+\begin{figure}
+\centering \includegraphics[width=3in]{pseudocode}
+\caption{Pseudocode from Aine et al., detailing the original single-threaded version of the algorithm we parallelized.}
+\label{fig:pseudocode}
 \end{figure}
+
+Our algorithm is based off of the single-threaded, round-robin SMHA* code presented in \cite{Aine14} (Figure \ref{fig:pseudocode}).
+Our parallelized code was written in C++ with inter-process communication accomplished using MPI.
+The work was distributed over one or more of the 16-core machines in the Apt cluster.
+% TODO: Machine specs here?
+Or in the experimental setup section?
+
+\subsection{Design Decisions}
+
+\textbf{Interprocess communication with MPI}
+
+We investigated other frameworks designed for iterative algorithms on a stable dataset such as Spark and GraphLab.
+However, most of these frameworks were designed for chunking up a large dataset upon which a single operation is performed.
+Our application's needs differed from this model in that the dataset being operated on (the graph being explored) was shared by all workers, and the workers themselves were performing differing operations on subsets of the same data.
+
+In the case where the working set of data is small enough to be contained within each process, the system is much less complex, and the services provided by Spark and GraphLab are not so crucial.
+For these reasons we instead decided to distribute our code using MPI.
+
+\textbf{Synchronous broadcast vs. asynchronous sync-ups}
+
+% TODO: Fill me in! MPI-specific feature...
+
+\textbf{Update collation performed at anchor node}
+
+In our system, updates are sent from heuristic machines to the anchor machine, which collects these updates and combines them into a smaller set of table updates.
+For example, if two heuristic machines both send new shortest path values for the same state, the anchor machine receives both and at a later time sends out an update on the shortest path to that state that is the minimum of values it received.
+All nodes then update their data structures with this new information.
+% TODO: ^ Is this right?
+
+One alternative to this method of communicating updates would be to have each node periodically broadcast its table updates to all other nodes and each node would incorporate those updates in its own table.
+This requires many more messages to go over the network and work to be duplicated.
+However, in the case where the anchor node is overloaded or has a possibility of going down, this alternative would be more appealing.
+We didn't consider this to be within the scope of this project, however, so we only implemented a system of update collation at the anchor machine.
+However, this may be interesting to investigate in the future.
+%TODO: Maybe justfiy or link with any result that indicates that the anchor node's workload becomes a bottleneck (or doesn't)
 
@@
 
-%\section{Goals}
+We tested our algorithm on a set of four random sliding tile puzzles of size 5x5, averaging the results of each trial into a single result.
+The data we collected were (1) running time, (2) number of states in the problem graph that were expanded during the course of searching for a solution, and (3) the length in steps of the best solution at termination.
 
-%\subsection{75\% Goal}
-%\begin{enumerate}
-%\item Implement parallel versions of IMHA* and SMHA*
-%\item Identify bottlenecks in performance in running SMHA*
-%\end{enumerate}
+%TODO: ^ Explain how the states expanded node is calculated? Maybe as a footnote?
 
-%\subsection{100\% Goal}
-%\begin{enumerate}
-%\item Implement distributed versions of IMHA* and SMHA*, perhaps with fancier distributed heuristic selection (there exists recent work which uses a simple form of online learning)
-%\end{enumerate}
+% TODO: Apt? Machine specs?
 
-%\subsection{125\% Goal}
-%\begin{enumerate}
-%\item Design and implement a more sophisticated intermediate between IMHA* and SMHA*, based on the analysis of parallel and distributed system tradeoffs from our preliminary evaluations.
-%\end{enumerate}
+% TODO: Talk about what we varied for the tests.
+Frequency of communication, effect of adding more heuristics, effect of adding more machines (maybe we want a graph that compares the effect of adding more heuristics on the same machine vs.
+adding more heuristics by using more machines), and comparison to single thread/round robing running time
 
-%\section{Resources Needed}
+In MHA*, two weights are specified.
+The first weight ($w_1$) determines how much priority to give heuristic expansions over the anchor search.
+A higher $w_1$ results in more depth-first-search-like behavior.
+The second weight ($w_2$) determines how close to the current known minimum path length a path to the goal must be before being selected as a solution.
+MHA* gives a guarantee that the solution found at termination is at most $w_1 * w_2$ times the optimum value.
+For all experiments, we used $w_1$, $w_2 = 2$.
+Note that when there is only one heuristic in a test (the admissible heuristic), the optimal shortest path to the goal will be found by only having to expand each state at most once (guaranteed by running A* with an admissible heuristic) and neither $w_1$ nor $w_2$ have any effect.
 
-%We will probably need a machine with at least 32 cores for the 75\% goal and a cluster of some number of machines for the distributed version (100\% goal).
-We will be using MPI for message-passing between computers and C++ to program.
+\section{Results}
 
-%\section{Relevance to the Class}
+% TODO: I need to put a ton of graphs and tables here
 
-%We will investigate ways to parallelize SMHA*, and compare its performance to parallelized IMHA*.
-This project gives us the opportunity to explore classic tradeoffs in designing a distributed system, particularly between network latency and local computation.
+\section{Conclusions}
 
-%\section{Work Distribution}
+% TODO: Fill me in!
 
-%Tasks for 75\% goal:
-%\begin{enumerate}
-%\item Implement IMHA* using locks / shared memory
-%\item Implement SMHA* using locks / shared memory
-%\item Pull problem and heuristics from cited work
-%\item Run algorithms on at least one of the cited example problems
-%\item Evaluate performance overhead of the parallelism
-%\end{enumerate}
-
-%Tasks for 100\% goal:
-%\begin{enumerate}
-%\item Implement IMHA*/SHMA* using message passing
-%\item Run algorithms on a cluster of machines
-%\item Evaluate performance overhead of local and networked message passing
-%\item Reflect and compare against multicore version
-%\end{enumerate}
-
-%Tasks for 125\% goal:
-%\begin{enumerate}
-%\item Implement new version of SMHA* based on considerations of the evaluation
-%\item Test and evaluate
-%\end{enumerate}
-%\newpage
+% TODO: Should we have some kind of discussion on whether the parallelized version of SMHA* keeps the w1*w2 optimality guarantees that the single-threaded version does??
+Also, maybe part of this, but the longer we go without doing a table update sync-up, the more the algorithm resembles IMHA*.
+It might be good to point this out somewhere, maybe in the Results/Discussion section
 
@@
 %\subsection{Introduction}
-
-%Similar to the above problem statement.
-
 %\subsection{Related Work}
-
-%\cite{Aine14} (MHA*) presents the original multi-heuristic anchored search.
-\cite{Phillips14} (parallelized A* for slow expansions) presents a criterion for parallel expansion using locks, while retaining the approximate suboptimality guarantees of sequential wA*.
-
 %\subsection{System Description}
-
-%A distributed system running one heuristic on each machine.
-
 %\subsection{Evaluation}
-
-%\textbf{IMHA* experiment:}
-
-%We will implement IMHA* on 32 cores, each running its own heuristic.
-
-%\textbf{SMHA* experiment:}
-
-%We will implement SMHA* on 32 cores, each running its own heuristic and updating shared data structures for path costs.
-
-%We want to keep the approximate optimality guarantees of unparallelized MHA*, so this requires careful maintenance of the data structures to ensure the invariants are kept.
-
-%Explore different way of updating the shared data (i.e. data replication and consistency checking/message-passing?)?
-
-%Performance of the various implementations will be evaluated on the problems from Aine et al.:
-%\begin{enumerate}
-%\item Robot arm manipulation
-%\item 3D path planning (navigation)
-%\item Sliding tile puzzle
-%\end{enumerate}
-
-\subsection{Conclusions and Future Work}
+%\subsection{Conclusions and Future Work}
 

### 2014-12-10 SMHA.cpp
lame is being LAAAAAAAAAME too fast

@@ -1,3 +1,4 @@
+// TODO: replace final broadcast - just send a normal message back to anchor
 #include <array>
 #include <cmath>
 #include <thread>
@@ -47,6 +48,16 @@ string filename = "test";
 
 // right, up, left, down
 constexpr array<int, NUM_MOVES> moves = { 1, -GRID_COLS, -1, GRID_COLS };
+
+Clock::time_point debug_t1, debug_t2;
+void debug_time(string str) {
+  debug_t1 = debug_t2;
+  debug_t2 = Clock::now();
+  
+  double dt = chrono::duration<double,chrono::seconds::period>(debug_t2-debug_t1).count();
+  if (dt > 0.0001)
+    cout << str << ' ' << dt << endl;
+}
 
 // print a nicely formatted board
 void printState(const State& s) {
@@ -395,7 +406,7 @@ public:
   }
 
   void run() {
-	  int benchmark = 100000;
+	  int benchmark = 0;
     int flag = 0;
     MPI_Status status;
     MPI_Request request;

### 2014-12-10 lame_SMHA.cpp

@@ -24,7 +24,7 @@ typedef chrono::high_resolution_clock Clock;
 typedef pair<vector<int>, vector<int> > State;
 
 // compile-time constants
-constexpr int NUM_HEURISTICS = 16;
+constexpr int NUM_HEURISTICS = 4;
 constexpr int MASK_CLOSED = 1;
 constexpr int MASK_CLOSED_ANCHOR = 2;
 constexpr int GRID_ROWS = 5;
@@ -42,6 +42,16 @@ string filename = "test";
 
 // right, up, left, down
 constexpr array<int, NUM_MOVES> moves = { 1, -GRID_COLS, -1, GRID_COLS };
+
+Clock::time_point debug_t1, debug_t2;
+void debug_time(string str) {
+  debug_t1 = debug_t2;
+  debug_t2 = Clock::now();
+  
+  double dt = chrono::duration<double,chrono::seconds::period>(debug_t2-debug_t1).count();
+  if (dt > 0.0001)
+    cout << str << ' ' << dt << endl;
+}
 
 // print a nicely formatted board
 void printState(const State& s) {
@@ -362,7 +372,7 @@ void prepareDistributedSearch() {
 }
 
 void runDistributedSearch() {
-  int benchmark = 100000;
+  int benchmark = 0;
   int flag = 0;
   int error = -1;
 
@@ -372,7 +382,7 @@ void runDistributedSearch() {
   time_elapsed = 0;
   // repeat until some thread declares the search to be finished
   while (finished == -1 && time_elapsed <= TIME_LIMIT) {
-    for (Searcher searcher : vec_search) {
+    for (Searcher& searcher : vec_search) {
       searcher.runSingleIteration();
     }
     // timing

### 2014-12-10 SMHA.cpp
SEED IN OUTER LOOP. GAASP

@@ -732,7 +732,8 @@ int main(int argc, char** argv) {
       cout << "Found path of length " << path_length << " Discovered nodes = " << searcher.num_discovered << ". Expanded nodes: " << searcher.num_expanded << ". Time: " << searcher.time_elapsed << endl;
 		}
 		cout << "Total discovered: " << searcher.total_discovered << " total expanded: " << searcher.total_expanded << endl;
-		fprintf(fout, "%f %f %f %d %f\n", searcher.w1, searcher.w2, searcher.time_elapsed, searcher.num_expanded, path_length);
+    fprintf(fout, "%f %f %f %d %f\n", searcher.w1, searcher.w2, searcher.time_elapsed, searcher.total_expanded, path_length);
+		//fprintf(fout, "%f %f %f %d %f\n", searcher.w1, searcher.w2, searcher.time_elapsed, searcher.num_expanded, path_length);
 	}
   }
   fclose(fout);

### 2014-12-10 lame_SMHA.cpp

@@ -24,7 +24,6 @@ typedef chrono::high_resolution_clock Clock;
 typedef pair<vector<int>, vector<int> > State;
 
 // compile-time constants
-constexpr int NUM_HEURISTICS = 4;
 constexpr int MASK_CLOSED = 1;
 constexpr int MASK_CLOSED_ANCHOR = 2;
 constexpr int GRID_ROWS = 5;
@@ -38,6 +37,7 @@ constexpr int HEAD_NODE = 0;
 
 static int finished = -1;
 int SEED = 1;
+int NUM_HEURISTICS = 1;
 string filename = "test";
 
 // right, up, left, down
@@ -198,8 +198,8 @@ void computeH(const State& s, StateData& s_data) {
   Cost hLC = linearConflicts(s, goal);
   Cost hMT = misplacedTiles(s, goal);
   for (int i = 0; i < NUM_HEURISTICS; ++i) {
-    s_data.h[i] = vec_search[i].MD*hMD + vec_search[i].LC*hLC + vec_search[i].MT*hMT;
-    s_data.iter[i] = vec_search[i].open.cend();
+    s_data.h.push_back(vec_search[i].MD*hMD + vec_search[i].LC*hLC + vec_search[i].MT*hMT);
+    s_data.iter.push_back(vec_search[i].open.cend());
   }
 }
 
@@ -406,6 +406,9 @@ int main(int argc, char** argv) {
     if (arg == "-filename" || arg == "-f") {
       filename = argv[++i];
     }
+    if (arg == "-numh" || arg == "-n") {
+      NUM_HEURISTICS = atoi(argv[++i]);
+    }
   }
 
   FILE* fout = fopen((filename + ".csv").c_str(), "a");
@@ -428,9 +431,9 @@ int main(int argc, char** argv) {
 		cout << "Total discovered: " << total_discovered << " total expanded: " << total_expanded << endl;
 
     // report stats
-    /*printf("map %d: Path Length=%f Visited Nodes=%d Explored Nodes=%d Planning Time=%f\n", i, path_length, searcher.num_discovered, searcher.num_expanded, time_elapsed);
-    fprintf(fout, "%f %f %f %d %f\n", w1, w2, time_elapsed, searcher.num_expanded, path_length);
-	cout << "Rank " << comm_rank << " Total discovered: " << total_discovered << " total expanded: " << total_expanded << endl;*/
[…]

### 2014-12-10 lame_SMHA.h

@@ -8,9 +8,9 @@ public:
   int mask;
   // g is the cost of a discovered path, h estimates the remaining cost to goal
   Cost g;
-  Cost h[NUM_HEURISTICS];
+  vector<Cost> h;
   // points to the state's location in the priority queue
-  multimap<Cost, State>::const_iterator iter[NUM_HEURISTICS];
+  vector<multimap<Cost, State>::const_iterator> iter;
   // initialization
   StateData() : mask(0), g(INFINITE) {}
 };

### 2014-12-10 test.sh

@@ -3,12 +3,16 @@
 #set eps, sweep time
 #4 16 32
 mpic++ -std=c++0x -o SMHA SMHA.cpp
+g++ -std=c++0x -o lame_SMHA lame_SMHA.cpp
 
-for NUM_PROC in 1 8
+
+for SEED in 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20
 do
-  rm stats_${NUM_PROC}.csv
-  for SEED in 1 2 3 4
+  rm stats_cool_${NUM_PROC}.csv
+  rm stats_lame_${NUM_PROC}.csv
+  for NUM_PROC in 1 2 4 8 16
   do
-    mpirun -n ${NUM_PROC} -hostfile host_file ./SMHA -seed ${SEED} -filename stats_${NUM_PROC}
+    mpirun -n ${NUM_PROC} -hostfile host_file ./SMHA -s ${SEED} -f stats_cool_${NUM_PROC}
+    ./lame_SMHA -n ${NUM_PROC} -s ${SEED} -f stats_lame_${NUM_PROC}
   done
 done

### 2014-12-10 SMHA.cpp
Added debug code to SMHA

@@ -537,6 +537,7 @@ public:
         // flush and kill, but don't kill the master!!!!!
 		  finished = 1;
 		  cout << "Machine " << comm_rank << " found a solution" << endl;
+		  cout << "Time taken: " << time_elapsed << ". Total expanded: " << total_expanded << " Local expanded: " << num_expanded << endl;
 
 		  //Now do one final comm to inform the others.
 		  if (comm_rank == HEAD_NODE) {

### 2014-12-10 paper.tex
intro writing stuff

@@
 Weighted A* is a simple, heuristic-based search algorithm used in artificial intelligence applications such as robotics and games.
-A heuristic function estimates the remaining distance from any state to the goal, thus guiding the search toward the goal.
+A heuristic function estimates the remaining distance from any state to the goal.
+Whereas a breadth-first method such as Dijkstra's algorithm will expand exponentially many states relative to the goal distance, wA* will focus toward the goal when guided by a good heuristic.
 If the heuristic is \textbf{admissible}, meaning it never overestimates the true distance, then the solution is guaranteed to be optimal to within the weight factor.
-If it meets a stronger condition called \textbf{admissibility}, then no state needs to be expanded more than once.
+If it meets a stronger condition called \textbf{monotonicity}, then no state needs to be expanded more than once.
 
-Finding good heuristic functions is difficult, and moreover, finding an admissible heuristic function that is reasonably accurate over the entirety of the search space is often not practical.
-Multi-heuristic A* (MHA*) \cite{Aine14} is an alteration of the A* algorithm that allows the usage of a combination of multiple arbitrary heuristics to guide the search, while retaining the optimality bound of wA* provided that one consistent heuristic is used as an “anchor”.
+Finding good efficiently computable heuristic functions is difficult and, moreover, finding an admissible heuristic function that is reasonably accurate over the entirety of the search space is often not practical.
+Multi-heuristic A* (MHA*) \cite{Aine14} is an alteration of the A* algorithm that allows combines multiple arbitrary heuristics to guide the search; to retain the optimality bound, it is only necessary to use one monotone heuristic, called an ``anchor".
 Where the shortcomings of each individual heuristic may result in local optima which trap the search, the ability to change heuristics can provide an escape.
@@
 In their work, Aine et al.
-propose two variants on MHA*: Independent Multi-Heuristic A* (IMHA*), where path cost and shortest path information are tracked separately for each heuristic, and Shared Multi-Heuristic A* (SMHA*), where shared knowledge of path cost and shortest path information are updated by all heuristics.
-SMHA* is the more effective of the two (explores the graph in a less repetitive manner and enables more co-operation between heuristics), but is more difficult to parallelize due to the amount of shared data.
+propose MHA* in two flavors: Independent Multi-Heuristic A* (IMHA*), where path cost and shortest path information are tracked separately for each heuristic, and Shared Multi-Heuristic A* (SMHA*), where shared knowledge of path cost and shortest path information are shared and updated by all heuristics.
+IMHA* would be almost trivial to parallelize, as it requires minimal communication between the searches run by different heuristics.
+SMHA* uses a very large amount of frequently updated shared memory, presenting major challenges for parallel settings, let alone distributed systems settings.
 
-For this project, we implemented a distributed version of this second variant, SMHA*.
+Nonetheless, as SMHA* is the more effective of the two (enabling more co-operation and less redundancy between heuristics), we decided for our project to implement a distributed version of SMHA*.
 
@@
 
-Our work is directly based off of Aine et al.'s original multi-heuristic anchored search \cite{Aine14} mentioned above.
-A parallelized version of the original A* algorithm was presented by Phillips et al.
-\cite{phillips2014pa}.
-This method of parallelization makes inferences on when it's safe to expand multiple nodes in parallel, and in so doing maintains the approximate suboptimality guarantees of sequential wA*.
-This parallelization is dependent on the expansion operation being expensive enough to justify expanding multiple nodes in parallel, which may not be the case for all problems.
-By contrast, our method parallelizes the search process by constantly expanding multiple nodes chosen by different heuristics (instead of multiple nodes chosen by a single heuristic).
-As a result, our method differs from the previous in that, although nodes may be expanded multiple times by different machines, we do not require expansions to be slow.
+Our work is directly based on Aine et al.'s original multi-heuristic A* search \cite{Aine14} mentioned above.
+There exists past work on the paralellization of A*; in particular, Phillips et al.
+\cite{phillips2014pa} were able to achieve parallelization while retaining wA*'s guarantee of suboptimality without reexpansions.
+This method of parallelization makes inferences on when it's safe to expand multiple nodes in parallel, by resolving dependencies between potential optimal paths.
+Their speed gains are dependent on the expansion operation being expensive enough to dominate the cost of computing these dependency checks, which may not be the case for all problems.
+By contrast, our method parallelizes the search process by simultaneously expanding nodes chosen by different heuristics, instead of multiple nodes chosen by a single heuristic.
+We allow each state to be expanded by each heuristic at most once and, furthermore, no state will be expanded by a non-anchor heuristic which has received a message from another heuristic after the latter has expanded the same state.
+In other words, aside from communication latency, no state will be expanded more than twice.
 

### 2014-12-10 lame_SMHA.cpp

@@ -26,8 +26,8 @@ typedef pair<vector<int>, vector<int> > State;
 // compile-time constants
 constexpr int MASK_CLOSED = 1;
 constexpr int MASK_CLOSED_ANCHOR = 2;
-constexpr int GRID_ROWS = 5;
-constexpr int GRID_COLS = 5;
+constexpr int GRID_ROWS = 4;
+constexpr int GRID_COLS = 4;
 constexpr int NUM_MOVES = 4;
 constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
@@ -317,8 +317,8 @@ vector<State> Searcher::getSolution() {
 
 
 void prepareDistributedSearch() {
-  w1 = 2.0;
-  w2 = 2.0;
+  w1 = 1.0;
+  w2 = 1.0;
   // make the standard orderly goal state
   for (int i = 1; i < GRID_ROWS*GRID_COLS; i++) {
     goal.first.push_back(i);

### 2014-12-11 paper.tex
Adding graphs and data table skeleton to results section

--- 2014-12-10 paper.tex
@@
 \usepackage{bold-extra}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 
-% TODO: I need to put a ton of graphs and tables here
+\begin{figure}
+\centering
+\begin{subfigure}{3.2in}
+  \includegraphics[width=3.2in]{graphs/singlethreadcomp_runningtime}
+  \caption{Running time}
+  \label{fig:singthread_runtime}
+\end{subfigure}
+\begin{subfigure}{3.2in}
+  \includegraphics[width=3.2in]{graphs/singlethreadcomp_solutionquality}
+  \caption{Solution length}
+  \label{fig:singthread_solnqual}
+\end{subfigure}
+\caption{Comparison with Runtime performance and solution quality as the number of different heuristics (cores) on a single machine increases.}
+\label{fig:singthread}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}{3.2in}
+  \includegraphics[width=3.2in]{graphs/numcores_runningtime}
+  \caption{Running time}
+  \label{fig:numcores_runtime}
+\end{subfigure}
+\begin{subfigure}{3.2in}
+  \includegraphics[width=3.2in]{graphs/numcores_solutionquality}
+  \caption{Solution length}
+  \label{fig:numcores_solnqual}
+\end{subfigure}
+\caption{Runtime performance and solution quality as the number of different heuristics (cores) on a single machine increases.}
+\label{fig:numcores}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}{3.2in}
+  \includegraphics[width=3.2in]{graphs/nummachines_runningtime}
+  \caption{Running time}
+  \label{fig:nummachines_runtime}
+\end{subfigure}
+\begin{subfigure}{3.2in}
+  \includegraphics[width=3.2in]{graphs/nummachines_solutionquality}
+  \caption{Solution length}
+  \label{fig:nummachines_solnqual}
+\end{subfigure}
+\caption{Runtime performance and solution quality as the number of different machines (with one process each) increases from 1 to 16.}
+\label{fig:nummachines}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}{2.0in}
+  \includegraphics[width=2.0in]{graphs/commfreq_runningtime}
+  \caption{Running time}
+  \label{fig:commfreq_runtime}
+\end{subfigure}
+\begin{subfigure}{2.0in}
+  \includegraphics[width=2.0in]{graphs/commfreq_expansions}
+  \caption{Number of state expansions}
+  \label{fig:commfreq_expansions}
+\end{subfigure}
+\begin{subfigure}{2.0in}
+  \includegraphics[width=2.0in]{graphs/commfreq_solutionquality}
+  \caption{Solution length}
+  \label{fig:nummachines_solnqual}
+\end{subfigure}
+\caption{Runtime performance, solution quality, and number of states expanded by the heuristic machines as the number of iterations between communication increases.
+*At the high end (12800 iterations between sync-ups), one of the four trials timed out (failed to find a solution within 5 minutes).
+The values in this graph are the average of the remaining three trials.}
+\label{fig:commfreq}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tabular}{ c | c | c | c }
+    \hline
+    \multicolumn{4}{c}{Single-threaded SMHA* implementation} \\ \hline    
+    \# cores & time & \# expand & soln len \\ \hline
+    1 & 1 & 2 & 3 \\ \hline
+    2 & 4 & 5 & 6 \\ \hline
+    4 & 7 & 8 & 9 \\ \hline
+    % FILL USING AMPERSAND-SEPARATED VALUES WITH "\\ \hline" APPENDED AT THE END
+
+     \multicolumn{4}{c}{Parallelized SMHA* (single machine, syncing every X iterations)} \\ \hline
+    \# cores & time & \# expand & soln len \\ \hline
+    % FILL USING AMPERSAND-SEPARATED VALUES WITH "\\ \hline" APPENDED AT THE END
+
+     \multicolumn{4}{c}{Distributed SMHA* (one process per machine, syncing every X iterations)} \\ \hline
+    \# machines & time & \# expand & soln len \\ \hline
+    % FILL USING AMPERSAND-SEPARATED VALUES WITH "\\ \hline" APPENDED AT THE END
+
+     \multicolumn{4}{c}{Distributed SMHA* with varying frequency of sync-ups} \\ \hline
+     \multicolumn{4}{c}{(4 processes across 2 machines)} \\ \hline
+    \# iterations & time & \# expand & soln len \\ \hline
+    % FILL USING AMPERSAND-SEPARATED VALUES WITH "\\ \hline" APPENDED AT THE END
+
+    \hline
+\end{tabular}
+\caption{Data from our trials}
+\label{fig:datatable}
+\end{figure}
 

### 2014-12-11 SMHA_Async.cpp
Updates to lame, SMHA, and Async

@@ -27,20 +27,18 @@ typedef pair<vector<int>, vector<int> > State;
 // compile-time constants
 constexpr int MASK_CLOSED = 1;
 constexpr int MASK_CLOSED_ANCHOR = 2;
-constexpr int GRID_ROWS = 6;
-constexpr int GRID_COLS = 6;
+constexpr int GRID_ROWS = 5;
+constexpr int GRID_COLS = 5;
 constexpr int NUM_MOVES = 4;
 constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
 
 //communication constants
-constexpr bool COMM_ITER = true;
-constexpr double COMM_INTERVAL = 0.2;
-constexpr int COMM_FREQ = 10000;
-constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*4; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
+constexpr int COMM_FREQ = 1000;
+constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*10; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
 constexpr int DATUM_SIZE = (GRID_ROWS * GRID_COLS) + 1 + 2; //4x4 state, backtrace, mask, g, and h. Assumes Cost is same size as int
 constexpr int HEAD_NODE = 0;
-constexpr int SLOWDOWN = 1;	//Factor to slow down heuristic calculation by. Must be an int and >= 1. Make 1 for no slowdown
+constexpr int SLOWDOWN = 10;	//Factor to slow down heuristic calculation by. Must be an int and >= 1. Make 1 for no slowdown
 
 static int comm_size;
 static int comm_rank;
@@ -244,8 +242,8 @@ public:
   int total_expanded;
 
   //Communication stuff
-  int * head_buffer;
-  int * child_buffer;
+  int * to_buffer;
+  int * from_buffer;
   //set<State> updated_states;
   unordered_set< State, state_hash> updated_states;
 
@@ -259,8 +257,8 @@ public:
 		  hLC = linearConflicts(s, goal);
 		  hMT = misplacedTiles(s, goal);
 	  }
-    s_data.hAnch = hMD + hLC;
-    s_data.h = MD*hMD + LC*hLC + MT*hMT;
+	s_data.hAnch = hMD + hLC;
[…]

### 2014-12-11 lame_SMHA.cpp

@@ -33,11 +33,13 @@ constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
 constexpr int HEAD_NODE = 0;
 
+constexpr int SLOWDOWN = 10;
+
 #include "lame_SMHA.h"
 
 static int finished = -1;
 int SEED = 1;
-int NUM_HEURISTICS = 1;
+int NUM_HEURISTICS = 8;
 string filename = "test";
 
 // right, up, left, down
@@ -194,9 +196,14 @@ vector<Searcher> vec_search;
 
 // determine the heuristics to be used
 void computeH(const State& s, StateData& s_data) {
-  Cost hMD = manhattanDist(s, goal);
-  Cost hLC = linearConflicts(s, goal);
-  Cost hMT = misplacedTiles(s, goal);
+	Cost hMD;
+	Cost hLC;
+	Cost hMT;
+	for (int i = 0; i < SLOWDOWN; i++) {
+		hMD = manhattanDist(s, goal);
+		hLC = linearConflicts(s, goal);
+		hMT = misplacedTiles(s, goal);
+	}
   for (int i = 0; i < NUM_HEURISTICS; ++i) {
     s_data.h.push_back(vec_search[i].MD*hMD + vec_search[i].LC*hLC + vec_search[i].MT*hMT);
     s_data.iter.push_back(vec_search[i].open.cend());

### 2014-12-11 paper.tex
hope it merged right

@@
 Each machine runs its own heuristic, with the anchor machine's being the admissible heuristic.
-Each of the non-anchor heuristic machines periodically send updates of the data structures (tables containing shortest path information found so far) to the anchor machine, which compares the data it receives and sends out table updates of the best paths known so far to all other nodes.
+Each of the non-anchor heuristic machines periodically sends updates of the data structures (tables containing shortest path information found so far) to the anchor machine, which compares the data it receives and sends out table updates of the best paths known so far to all other nodes.
 
@@
 
-In our system, updates are sent from heuristic machines to the anchor machine, which collects these updates and combines them into a smaller set of table updates.
-For example, if two heuristic machines both send new shortest path values for the same state, the anchor machine receives both and at a later time sends out an update on the shortest path to that state that is the minimum of values it received.
-All nodes then update their data structures with this new information.
-% TODO: ^ Is this right?
+\textbf{Update collation performed at anchor node}
 
-One alternative to this method of communicating updates would be to have each node periodically broadcast its table updates to all other nodes and each node would incorporate those updates in its own table.
-This requires many more messages to go over the network and work to be duplicated.
-However, in the case where the anchor node is overloaded or has a possibility of going down, this alternative would be more appealing.
-We didn't consider this to be within the scope of this project, however, so we only implemented a system of update collation at the anchor machine.
-However, this may be interesting to investigate in the future.
-%TODO: Maybe justfiy or link with any result that indicates that the anchor node's workload becomes a bottleneck (or doesn't)
+SMHA* needs to share the values of $g$ (distance estimates) and $bp$ (back-pointers) for every state.
+As sharing memory is infeasible in our distributed implementation, processes periodically broadcast their $g$ and $bp$ values to update one another.
+This can result in conflicting data for a particular state, which is always resolved by taking the data which gives the lowest $g$ value for that state as it contains the best estimate found so far.
+During conflict resolution, we also take the inclusive OR of the flags specifying whether the state has ever been expanded by the anchor, or by any machine at all.
+
+Each broadcast by a heuristic machine only needs to contain the portion of the graph which has been updated by the machine since its most recent network communication.
+Nonetheless, if every machine broadcasts its updates directly to all other machines, the network would flood with messages and much conflict-resolution work would be duplicated.
+Therefore, we have every machine send its table updates exclusively to the anchor machine.
+The anchor alone takes the burden of resolving all conflicts between duplicate states, and then periodically broadcasts the merged collection of updates to all machines.
+
+The disadvantage is that we rely heavily on the anchor machine functioning reliably.
+This is a small price to pay, given that the anchor is already necessary for updating the optimality bound.
+To avoid unduly burdening the anchor, every machine is able to compute the anchor heuristic and includes its value in the state updates it sends to the anchor machine.
+
+\textbf{Back-pointer optimization}
+
+$bp(s)$ is a pointer to the parent of $s$ along the optimal path found so far from the start to $s$.
+It is not needed to perform the search and compute a conservative distance to the goal; it is only used to reconstruct the path once the search terminates.
 

### 2014-12-11 SMHA.cpp

@@ -27,8 +27,8 @@ typedef pair<vector<int>, vector<int> > State;
 // compile-time constants
 constexpr int MASK_CLOSED = 1;
 constexpr int MASK_CLOSED_ANCHOR = 2;
-constexpr int GRID_ROWS = 4;
-constexpr int GRID_COLS = 4;
+constexpr int GRID_ROWS = 5;
+constexpr int GRID_COLS = 5;
 constexpr int NUM_MOVES = 4;
 constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
@@ -705,8 +705,8 @@ public:
 Searcher searcher;
 
 void prepareDistributedSearch() {
-  searcher.w1 = 1.0;
-  searcher.w2 = 1.0;
+  searcher.w1 = 2.0;
+  searcher.w2 = 2.0;
   // make the standard orderly goal state
   for (int i = 1; i < GRID_ROWS*GRID_COLS; i++) {
     searcher.goal.first.push_back(i);

### 2014-12-11 SMHA.cpp
SMHA changes

@@ -27,8 +27,8 @@ typedef pair<vector<int>, vector<int> > State;
 // compile-time constants
 constexpr int MASK_CLOSED = 1;
 constexpr int MASK_CLOSED_ANCHOR = 2;
-constexpr int GRID_ROWS = 4;
-constexpr int GRID_COLS = 4;
+constexpr int GRID_ROWS = 6;
+constexpr int GRID_COLS = 6;
 constexpr int NUM_MOVES = 4;
 constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
@@ -36,7 +36,7 @@ constexpr Cost INFINITE = 1e30;
 //communication constants
 constexpr bool COMM_ITER = true;
 constexpr double COMM_INTERVAL = 0.05;
-constexpr int COMM_FREQ = 100000;
+constexpr int COMM_FREQ = 1000;
 constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*4; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
 constexpr int DATUM_SIZE = (GRID_ROWS * GRID_COLS) + 1 + 2; //4x4 state, backtrace, mask, g, and h. Assumes Cost is same size as int
 constexpr int HEAD_NODE = 0;
@@ -705,8 +705,8 @@ public:
 Searcher searcher;
 
 void prepareDistributedSearch() {
-  searcher.w1 = 1.0;
-  searcher.w2 = 1.0;
+  searcher.w1 = 2.0;
+  searcher.w2 = 2.0;
   // make the standard orderly goal state
   for (int i = 1; i < GRID_ROWS*GRID_COLS; i++) {
     searcher.goal.first.push_back(i);

### 2014-12-11 graph_file.m
Adding graphing code, and .csv files needed to demonstrate how to use it

@@ -1,0 +1,203 @@
+
+function f = graph_file()
+
+labRuntime = 'Running time (seconds)';
+labExpansions = 'Number of expanded states';
+labSolution = 'Shortest solution';
+
+
+
+% EXAMPLE #1 OF HOW TO GRAPH STUFF
+filegroup = 'lame';
+xvals = [1, 2, 4, 8, 16];
+xlab = 'Number of nodes/heuristics (one machine)';
+[runtime, expansions, solution, error_r, error_e, error_s] = get_data(filegroup, xvals);
+
+graph_with_error_bars(xvals, runtime, error_r, xlab, labRuntime, true, true);
+graph_with_error_bars(xvals, expansions, error_e, xlab, labExpansions, true, true);
+graph_with_error_bars(xvals, solution, error_s, xlab, labSolution, true, false);
+% IMPORTANT!! FOR EACH GRAPH THAT POPS UP, YOU NEED TO ADD A TITLE MANUALLY
+%             AND THEN SAVE IT AS A PNG MANUALLY
+% /EXAMPLE #1
+
+
+
+% EXAMPLE #2 OF HOW TO GRAPH STUFF
+xvals = [1, 2, 4, 8, 16];
+xlab = 'Number of nodes/heuristics (one machine)';
+filegroup = 'lame';
+[runtime1, expansions1, solution1, error_r1, error_e1, error_s1] = get_data(filegroup, xvals);
+legendlab1 = 'Lame SMHA*';
+
+filegroup = 'cool';
+[runtime2, expansions2, solution2, error_r2, error_e2, error_s2] = get_data(filegroup, xvals);
+legendlab2 = 'Cool SMHA*';
+
+multigraph_with_error_bars(xvals, runtime1, runtime2, error_r1, error_r2, xlab, labRuntime, legendlab1, legendlab2, true, true);
+multigraph_with_error_bars(xvals, expansions1, expansions2, error_e1, error_e2, xlab, labExpansions, legendlab1, legendlab2, true, true);
+multigraph_with_error_bars(xvals, solution1, solution2, error_s1, error_s2, xlab, labSolution, legendlab1, legendlab2, true, false);
+% IMPORTANT!! FOR EACH GRAPH THAT POPS UP, YOU NEED TO ADD A TITLE MANUALLY
[…]

### 2014-12-11 paper.tex
backpointer talk

@@
 
-\titleformat{\section}{\normalsize\bf}{\thesection \hspace{5pt}}{0em}{}
-\titleformat{\subsection}{\normalsize\bf}{\thesubsection \hspace{5pt}}{0em}{}
+%\titleformat{\section}{\normalsize\bf}{\thesection \hspace{5pt}}{0em}{}
+%\titleformat{\subsection}{\normalsize\bf}{\thesubsection \hspace{5pt}}{0em}{}
 
@@
 
-\centerline{\large{\textbf{Distributed Multi-Heuristic A* (HAMSTAR)}}}
+\centerline{\Large{\textbf{Distributed Multi-Heuristic A* (HAMSTAR)}}}
 \centerline{Noam Brown, Aram Ebtekar, Yuzuko Nakamura}
@@
 
-\textbf{Interprocess communication with MPI}
+\subsubsection*{Interprocess communication with MPI}
 
@@
 
-\textbf{Synchronous broadcast vs. asynchronous sync-ups}
+\subsubsection*{Synchronous broadcast vs. asynchronous sync-ups}
 
@@
 
-\textbf{Update collation performed at anchor node}
-
-\textbf{Update collation performed at anchor node}
+\subsubsection*{Update collation performed at anchor node}
 
@@
 
-\textbf{Back-pointer optimization}
+\subsubsection*{Back-pointer optimization}
 
-$bp(s)$ is a pointer to the parent of $s$ along the optimal path found so far from the start to $s$.
-It is not needed to perform the search and compute a conservative distance to the goal; it is only used to reconstruct the path once the search terminates.
+$bp(s)$ represents the parent of $s$ along the optimal path found so far from the start to $s$.
+As a single pointer cannot be shared across a distributed replicated data structure, our first implementation used a full state representation (5x5 sliding tile puzzle) in place of $bp(s)$, nearly doubling the length of network messages.
+However, notice that $bp(s)$ is not needed to perform the search and compute $g$-values; it is only needed to reconstruct the solution path once the search terminates.
+
+Therefore, it suffices to use pointers referring to the machine's local entry for that state.
+These pointers need not be communicated until the search terminates.
+At that point, the path can be obtained by walking backwards from the goal.
+At each step, we move from the current state $s$ to the neighbor $s'$ which minimizes $g(s') + c(s',s)$ among the possible parents $s'=bp(s)$ pooled from all machines.
+Although we still end up sending full state representations, these are only for the parents of states along the path.
 

### 2014-12-11 test.sh

@@ -2,8 +2,8 @@
 
 #set eps, sweep time
 #4 16 32
-mpic++ -std=c++0x -o SMHA SMHA.cpp
-g++ -std=c++0x -o lame_SMHA lame_SMHA.cpp
+mpic++ -std=c++0x -O3 -o SMHA SMHA.cpp
+g++ -std=c++0x -O3 -o lame_SMHA lame_SMHA.cpp
 
 
 for SEED in 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20

### 2014-12-11 graph_file.m
Generated frequency graphs and matlab script fixes

@@ -8,38 +8,72 @@ labSolution = 'Shortest solution';
 
 
 % EXAMPLE #1 OF HOW TO GRAPH STUFF
+% Change the following if necessary
 filegroup = 'lame';
 xvals = [1, 2, 4, 8, 16];
 xlab = 'Number of nodes/heuristics (one machine)';
+titleRuntime = 'Running time with varying number of nodes';
+titleExpansions = 'States expanded with varying number of nodes';
+titleSolution = 'Solution length with varying number of nodes';
+
+% No need to change anything from this point on
 [runtime, expansions, solution, error_r, error_e, error_s] = get_data(filegroup, xvals);
 
-graph_with_error_bars(xvals, runtime, error_r, xlab, labRuntime, true, true);
-graph_with_error_bars(xvals, expansions, error_e, xlab, labExpansions, true, true);
-graph_with_error_bars(xvals, solution, error_s, xlab, labSolution, true, false);
-% IMPORTANT!! FOR EACH GRAPH THAT POPS UP, YOU NEED TO ADD A TITLE MANUALLY
-%             AND THEN SAVE IT AS A PNG MANUALLY
+fig_handle = graph_with_error_bars(xvals, runtime, error_r, xlab, labRuntime, true, true, titleRuntime);
+print(fig_handle, ['graphs/', filegroup, '_runningtime.png'], '-dpng', '-r0');
+fig_handle = graph_with_error_bars(xvals, expansions, error_e, xlab, labExpansions, true, true, titleExpansions);
+print(fig_handle, ['graphs/', filegroup, '_expansions.png'], '-dpng', '-r0');
+fig_handle = graph_with_error_bars(xvals, solution, error_s, xlab, labSolution, true, false, titleSolution);
+print(fig_handle, ['graphs/', filegroup, '_solutionquality.png'], '-dpng', '-r0');
 % /EXAMPLE #1
 
 
 
 % EXAMPLE #2 OF HOW TO GRAPH STUFF
+% Change the following if necessary
+filegroup1 = 'lame';
+filegroup2 = 'cool';
 xvals = [1, 2, 4, 8, 16];
 xlab = 'Number of nodes/heuristics (one machine)';
-filegroup = 'lame';
-[runtime1, expansions1, solution1, error_r1, error_e1, error_s1] = get_data(filegroup, xvals);
+titleRuntime = 'Running time with varying number of nodes';
[…]

### 2014-12-11 SMHA.cpp
Stats_cores10k data files

@@ -34,9 +34,9 @@ constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
 
 //communication constants
-constexpr bool COMM_ITER = true;
+constexpr bool COMM_ITER = false;
 constexpr double COMM_INTERVAL = 0.05;
-constexpr int COMM_FREQ = 1000;
+constexpr int COMM_FREQ = 100000;
 constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*4; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
 constexpr int DATUM_SIZE = (GRID_ROWS * GRID_COLS) + 1 + 2; //4x4 state, backtrace, mask, g, and h. Assumes Cost is same size as int
 constexpr int HEAD_NODE = 0;
@@ -534,6 +534,7 @@ public:
         //cout << "Process " << comm_rank << " Sending bcast" << endl;
 		Clock::time_point before_time = Clock::now();
         MPI_Bcast(child_buffer, (BUFFER_SIZE*DATUM_SIZE + 5)*comm_size, MPI_INT, HEAD_NODE, MPI_COMM_WORLD);
+		last_time = Clock::now();
 		time_waiting_bcast += chrono::duration<double, chrono::seconds::period>(Clock::now() - before_time).count();
         //cout << "Process " << comm_rank << " sent bcast " << endl;
         if (comm_rank != HEAD_NODE) {

### 2014-12-11 test_cores.sh
More cores data!

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+mpic++ -std=c++0x -o SMHA SMHA.cpp -O3
+
+for NUM_PROC in 2 4
+do
+  rm stats_cores10k_${NUM_PROC}.csv
+  for SEED in 1 2 3 4 5 6 7 8 9 10
+  do
+    mpirun -n ${NUM_PROC} -hostfile host_file16 ./SMHA -s ${SEED} -f stats_cores10k_${NUM_PROC}
+  done
+done

### 2014-12-11 test_lame.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+g++ -std=c++0x -o lame_SMHA lame_SMHA.cpp -O3
+
+
+for NUM_PROC in 32
+do
+  rm stats_lame_${NUM_PROC}.csv
+  for SEED in 1 2 3 4 5 6 7 8 9 10
+  do
+    ./lame_SMHA -n ${NUM_PROC} -s ${SEED} -f stats_lame_${NUM_PROC}
+  done
+done

### 2014-12-11 paper.tex
papery

@@
 
-We tested our algorithm on a set of four random sliding tile puzzles of size 5x5, averaging the results of each trial into a single result.
-The data we collected were (1) running time, (2) number of states in the problem graph that were expanded during the course of searching for a solution, and (3) the length in steps of the best solution at termination.
-
-%TODO: ^ Explain how the states expanded node is calculated? Maybe as a footnote?
+We tested our algorithm on the 24-puzzle, a 5x5 grid of sliding blocks which must be rearranged to achieve a standard goal configuration, somewhat akin to solving a Rubix cube.
+Uniformly at random, we generated ten starting configurations among the over $3\times 10^{23}$ states of the puzzle.
+The data we collected were (1) running time, (2) total number of states in the problem graph that were expanded across all processes during the search, and (3) the length in steps of the best solution at termination.
+Our main evaluation criterion is the time to complete the search with the required suboptimality guarantee, which we fix to $w_1 w_2 = 4$.
 
@@
 
-In MHA*, two weights are specified.
-The first weight ($w_1$) determines how much priority to give heuristic expansions over the anchor search.
-A higher $w_1$ results in more depth-first-search-like behavior.
-The second weight ($w_2$) determines how close to the current known minimum path length a path to the goal must be before being selected as a solution.
-MHA* gives a guarantee that the solution found at termination is at most $w_1 * w_2$ times the optimum value.
-For all experiments, we used $w_1$, $w_2 = 2$.
-Note that when there is only one heuristic in a test (the admissible heuristic), the optimal shortest path to the goal will be found by only having to expand each state at most once (guaranteed by running A* with an admissible heuristic) and neither $w_1$ nor $w_2$ have any effect.
+SMHA* specifies two weight parameters.
+The first weight ($w_1$) gives the search a more depth-first goal-directed focus.
+The second weight ($w_2$) determines the margin by which a proposed solution's cost may exceed the anchor's minimum frontier key value, which is itself $w1$ times a lower bound on the optimum solution cost.
+Therefore, SMHA* guarantees that the solution found at termination is at most $w_1 w_2$ times the optimum value.
+For all experiments, we used $w_1 = w_2 = 2$.
+Every state expanded by the anchor has a $g$ value which is optimal to within a factor of $w_1$.
 

### 2014-12-11 graph_file.m

@@ -9,7 +9,7 @@ labSolution = 'Shortest solution';
 
 % EXAMPLE #1 OF HOW TO GRAPH STUFF
 % Change the following if necessary
-filegroup = 'lame';
+filegroup = 'cores';
 xvals = [1, 2, 4, 8, 16];
 xlab = 'Number of nodes/heuristics (one machine)';
 titleRuntime = 'Running time with varying number of nodes';
@@ -32,14 +32,14 @@ print(fig_handle, ['graphs/', filegroup, '_solutionquality.png'], '-dpng', '-r0'
 % EXAMPLE #2 OF HOW TO GRAPH STUFF
 % Change the following if necessary
 filegroup1 = 'lame';
-filegroup2 = 'cool';
+filegroup2 = 'cores';
 xvals = [1, 2, 4, 8, 16];
 xlab = 'Number of nodes/heuristics (one machine)';
-titleRuntime = 'Running time with varying number of nodes';
-titleExpansions = 'States expanded with varying number of nodes';
-titleSolution = 'Solution length with varying number of nodes';
-legendlab1 = 'Lame SMHA*';
-legendlab2 = 'Cool SMHA*';
+titleRuntime = 'Running time with varying number of heuristics';
+titleExpansions = 'States expanded with varying number of heuristics';
+titleSolution = 'Solution length with varying number of heuristics';
+legendlab1 = 'Sequential';
+legendlab2 = 'Multi-core';
 
 % Do not need to change the following (I think)
 [runtime1, expansions1, solution1, error_r1, error_e1, error_s1] = get_data(filegroup1, xvals);
@@ -105,11 +105,11 @@ for i = 1:numfiles
   s_sort = sort(M(:, 5));
 
   runtime(i) = mean(r_sort(2:end-1)); % average values over trials and store
-  error_r(:,i) = [r_sort(1); r_sort(end)]; % save lowest and highest for error bars
+  error_r(:,i) = [r_sort(2); r_sort(end-1)]; % save lowest and highest for error bars
   expansions(i) = mean(e_sort(2:end-1));
-  error_e(:,i) = [e_sort(1); e_sort(end)];
+  error_e(:,i) = [e_sort(2); e_sort(end-1)];
   solution(i) = mean(s_sort(2:end-1));
-  error_s(:,i) = [s_sort(1); s_sort(end)];
+  error_s(:,i) = [s_sort(2); s_sort(end-1)];
 end
 
 end

### 2014-12-11 SMHA_Async.cpp
Even more data!

@@ -34,7 +34,7 @@ constexpr double TIME_LIMIT = 300;
 constexpr Cost INFINITE = 1e30;
 
 //communication constants
-constexpr int COMM_FREQ = 1000;
+constexpr int COMM_FREQ = 5000;
 constexpr int BUFFER_SIZE = COMM_FREQ*NUM_MOVES*10; //Number of new g values before a message is sent. The actual buffer is double this, because it needs the node too.
 constexpr int DATUM_SIZE = (GRID_ROWS * GRID_COLS) + 1 + 2; //4x4 state, backtrace, mask, g, and h. Assumes Cost is same size as int
 constexpr int HEAD_NODE = 0;
@@ -591,6 +591,9 @@ void run() {
       if (data[goal].g <= w2 * opt_bound) {
         // flush and kill, but don't kill the master!!!!!
 		  finished = 1;
+		  Cost path_length = data[goal].g;
+		  FILE* fout = fopen((filename + ".csv").c_str(), "a");
+		  fprintf(fout, "%f %f %f %d %f\n", w1, w2, time_elapsed, total_expanded, path_length);
 		  cout << "Machine " << comm_rank << " found a solution" << endl;
 		  cout << "Time taken: " << time_elapsed << ". Total expanded: " << total_expanded << " Local expanded: " << num_expanded << endl;
 		  cout << "Time commed: " << time_comm << endl;

### 2014-12-11 paper.tex
Added MPI section

@@
 Finding good efficiently computable heuristic functions is difficult and, moreover, finding an admissible heuristic function that is reasonably accurate over the entirety of the search space is often not practical.
-Multi-heuristic A* (MHA*) \cite{Aine14} is an alteration of the A* algorithm that allows combines multiple arbitrary heuristics to guide the search; to retain the optimality bound, it is only necessary to use one monotone heuristic, called an ``anchor".
+Multi-heuristic A* (MHA*) \cite{Aine14} is an alteration of the A* algorithm that allows multiple arbitrary heuristics to guide the search; to retain the optimality bound, it is only necessary to use one monotone heuristic, called an ``anchor".
 Where the shortcomings of each individual heuristic may result in local optima which trap the search, the ability to change heuristics can provide an escape.
@@
 \cite{phillips2014pa} were able to achieve parallelization while retaining wA*'s guarantee of suboptimality without reexpansions.
-This method of parallelization makes inferences on when it's safe to expand multiple nodes in parallel, by resolving dependencies between potential optimal paths.
+This method of parallelization makes inferences on when it is safe to expand multiple nodes in parallel, by resolving dependencies between potential optimal paths.
 Their speed gains are dependent on the expansion operation being expensive enough to dominate the cost of computing these dependency checks, which may not be the case for all problems.
@@
 
-In the case where the working set of data is small enough to be contained within each process, the system is much less complex, and the services provided by Spark and GraphLab are not so crucial.
-For these reasons we instead decided to distribute our code using MPI.
+We chose MPI, despite is notorious complexity, because our algorithm requires extremely fast and efficient communication between the anchor and the child machines.
+Moreover, since the entire state must be replicated across all machines, the necessary communication was relatively straightforward.
+Finally, we found detailed documentation on MPI, explaining the low-level workings of each command, much more readily accessible than comparable documentation on Spark or GraphLab, which seemed more focused on abstracting the details of the system away from the programmer.
+Given our emphasis on speed, MPI seemed like an appropriate choice.
+We ultimately used the OpenMPI implementation of the MPI standard.
 
-\subsubsection*{Synchronous broadcast vs. asynchronous sync-ups}
+\subsubsection*{Synchronous broadcast vs. one-on-one communication}
 
+We considered two communication models for our algorithm.
+The first one has the anchor periodically send each child a table of states.
+The table contains information for new states that were either received by the anchor, or calculated by the anchor, since the last time the anchor communicated with this child.
+Each child frequently checks for incoming messages using the \emph{MPI\_Iprobe} command; if a message is detected, the child receives the incoming data using \emph{MPI\_Recv} and responds with its own table of states.
+The anchor sends messages to each child one at a time, moving onto the next child when a response is received.
+
+The second communication model involves a period broadcast using the \emph{MPI\_Bcast} command.
+After a pre-determined number of iterations, all child machines send their new states to the anchor using \emph{MPI\_Gather}.
+The anchor sifts through these responses, and keeps only the best g value for each state.
+Every state that has had a new best g value found is then broadcast to all the child nodes.
+
+The one-on-one communication model has the benefit that it does not require all machines to cease work while communication happens; only the anchor and the single child machine need to participate in the communication.
+The downside is that the anchor must then do $O(N)$ communications, where $N$ is the number of children, and the time it takes to propogate the new state tables across the entire system is also $O(N)$.
+Broadcasts, on the other hand, require a simultaneous call from all machines.
+This generally requires more overhead, because it is unlikely that all machines will call the command at precisely the same time.
+However, once called, the Broadcast command allows the anchor to only send the data to a few machines in the network.
+These machines, in turn, forward the data to other nodes in the machine, and so on.
+Thus, the anchor does not spend any more time communicating than any other node in the system, and the new state table is propagated in $O(\log(N))$ time.
 % TODO: Fill me in! MPI-specific feature...

### 2014-12-11 paper.tex
graphs and heuristics

@@
 
+The anchor heuristic is $h(s) = MD(s) + LC(s)$ where $MD(s)$ is the sum of Manhattan distances of each non-blank tile to its destination, and $LC(s)$ is the linear conflicts heuristic which essentially adds 2 for each pair of horizontally or vertically aligned times which need to pass each other.
+This heuristic is admissible and monotone.
+Additional heuristics are generated randomly from $MT$, $LC$, and the misplaced tiles heuristic $MT$ which simply counts the number of non-blank tiles which are not already at their destinations.
+We randomly generate a triple of coefficients, each between 1 and 5, to form a linear combination of these heuristics.
+We repeat this process as many times as we require non-anchor heuristics.
+These heuristics are not at all admissible, and their high values give the search a goal-directed focus while the anchor is underestimating the true distance by a large margin.
+
 % TODO: Apt? Machine specs?
@@
 \begin{subfigure}{3.2in}
-  \includegraphics[width=3.2in]{graphs/singlethreadcomp_runningtime}
+  \includegraphics[width=3.2in]{graphs/figure1}
   \caption{Running time}
@@
 \begin{subfigure}{3.2in}
-  \includegraphics[width=3.2in]{graphs/singlethreadcomp_solutionquality}
+  \includegraphics[width=3.2in]{graphs/figure2}
   \caption{Solution length}
@@
 \begin{subfigure}{3.2in}
-  \includegraphics[width=3.2in]{graphs/numcores_runningtime}
+  \includegraphics[width=3.2in]{graphs/figure3}
   \caption{Running time}

### 2014-12-11 graph_file.m

@@ -9,7 +9,7 @@ labSolution = 'Shortest solution';
 
 % EXAMPLE #1 OF HOW TO GRAPH STUFF
 % Change the following if necessary
-filegroup = 'cores';
+filegroup = 'lame';
 xvals = [1, 2, 4, 8, 16];
 xlab = 'Number of nodes/heuristics (one machine)';
 titleRuntime = 'Running time with varying number of nodes';
@@ -32,14 +32,14 @@ print(fig_handle, ['graphs/', filegroup, '_solutionquality.png'], '-dpng', '-r0'
 % EXAMPLE #2 OF HOW TO GRAPH STUFF
 % Change the following if necessary
 filegroup1 = 'lame';
-filegroup2 = 'cores';
+filegroup2 = 'cool';
 xvals = [1, 2, 4, 8, 16];
 xlab = 'Number of nodes/heuristics (one machine)';
-titleRuntime = 'Running time with varying number of heuristics';
-titleExpansions = 'States expanded with varying number of heuristics';
-titleSolution = 'Solution length with varying number of heuristics';
-legendlab1 = 'Sequential';
-legendlab2 = 'Multi-core';
+titleRuntime = 'Running time with varying number of nodes';
+titleExpansions = 'States expanded with varying number of nodes';
+titleSolution = 'Solution length with varying number of nodes';
+legendlab1 = 'Lame SMHA*';
+legendlab2 = 'Cool SMHA*';
 
 % Do not need to change the following (I think)
 [runtime1, expansions1, solution1, error_r1, error_e1, error_s1] = get_data(filegroup1, xvals);
@@ -56,7 +56,8 @@ print(fig_handle, ['graphs/', filegroup1, filegroup2, 'comp_solutionquality.png'
 
 
 
-% FREQUENCY
+
+% EXAMPLE #3 FREQUENCY
 filegroup = 'freq';
 xvals = [100, 200, 400, 800, 1600, 3200, 6400, 12800];
 xlab = 'Number of iterations between sync-ups (4 nodes on 2 machines)';
@@ -73,7 +74,7 @@ fig_handle = graph_with_error_bars(xvals, expansions, error_e, xlab, labExpansio
 print(fig_handle, ['graphs/', filegroup, '_expansions.png'], '-dpng', '-r0');
[…]

### 2014-12-11 paper.tex
Abstract, big margins and size 11 font

@@
-\documentclass{article}
+\documentclass[11pt]{article}
 
@@
 
-\section{Abstract}
-
-A* is a popular graph search algorithm used in many artificial intelligence applications.
+\abstract{A* is a popular graph search algorithm used in many artificial intelligence applications.
 The original A* algorithm makes use of a single heuristic to prioritize exploration through a graph, but a variation on the algorithm known as multi-heuristic A* (MHA*) makes use of multiple heuristics to more quickly arrive at approximate solutions.
@@
 We show...
-[performance]
+[performance]}
 %TODO: Highlight important results
@@
 This heuristic is admissible and monotone.
-Additional heuristics are generated randomly from $MT$, $LC$, and the misplaced tiles heuristic $MT$ which simply counts the number of non-blank tiles which are not already at their destinations.
+Additional heuristics are generated randomly from $MD$, $LC$, and the misplaced tiles heuristic $MT$ which simply counts the number of non-blank tiles which are not already at their destinations.
 We randomly generate a triple of coefficients, each between 1 and 5, to form a linear combination of these heuristics.

### 2014-12-11 paper.tex
title
fourth graph to shatter the triforce
Added results

@@
 %\titleformat{\subsection}{\normalsize\bf}{\thesubsection \hspace{5pt}}{0em}{}
+\title{Distributed Multi-Heuristic A* (HAMSTAR)}
+\author{Noam Brown, Aram Ebtekar, Yuzuko Nakamura
+\\15-712 Fall 2014 Final Project Report}
 
 \begin{document}
-
-\centerline{\Large{\textbf{Distributed Multi-Heuristic A* (HAMSTAR)}}}
-\centerline{Noam Brown, Aram Ebtekar, Yuzuko Nakamura}
-\centerline{15-712 Fall 2014 Final Project Report}
+\maketitle
 
@@
 
+Our first set of experiments shows the performance of our algorithm as we increase the number of processes in the distributed setting.
+We consider the case of multiple processes on a single multi-core machine, as well as multiple machines each running a single process.
+For comparison, we also include the performance of the single-threaded SMHA* algorithm that our work is based on.
+For that algorithm, we show performance as the number of heuristics it uses increases.
+As shown, performance initially increases significantly, but then drops off as the number of processes increases.
+Moreover, our distributed algorithm performs comparably to SMHA* in the case where the number of heuristi
+
 \begin{figure}

### 2014-12-11 paper.tex
figure labels
prefer the terms sequential and monotone over single-threaded and consistent

@@
 \centering \includegraphics[width=5.0in]{system-diagram}
-\caption{A diagram of our system. One machine starts up the others and runs the admissible anchor search. The other machines run their individual heuristics. Communication is done using MPI. 
+\caption{A diagram of our system. One machine starts up the others and runs the monotone anchor search. The other machines run their individual heuristics. Communication is done using MPI. 
 \textcircled{1}\textbf{Anchor:} Starts heuristic machines; performs state expansions using anchor heuristic; receives, resolves, and broadcasts table updates; checks for termination conditions.
@@
 The main (anchor) machine starts up other heuristic machines.
-Each machine runs its own heuristic, with the anchor machine's being the admissible heuristic.
+Each machine runs its own heuristic, with the anchor machine's being the monotone heuristic.
 Each of the non-anchor heuristic machines periodically sends updates of the data structures (tables containing shortest path information found so far) to the anchor machine, which compares the data it receives and sends out table updates of the best paths known so far to all other nodes.
@@
 \centering \includegraphics[width=3in]{pseudocode}
-\caption{Pseudocode from Aine et al., detailing the original single-threaded version of the algorithm we parallelized.}
+\caption{Pseudocode from Aine et al., detailing the original sequential version of the algorithm we parallelized.}
 \label{fig:pseudocode}
@@
 
-Our algorithm is based off of the single-threaded, round-robin SMHA* code presented in \cite{Aine14} (Figure \ref{fig:pseudocode}).
+Our algorithm is based off of the sequential, round-robin SMHA* code presented in \cite{Aine14} (Figure \ref{fig:pseudocode}).
 Our parallelized code was written in C++ with inter-process communication accomplished using MPI.
@@
 The anchor heuristic is $h(s) = MD(s) + LC(s)$ where $MD(s)$ is the sum of Manhattan distances of each non-blank tile to its destination, and $LC(s)$ is the linear conflicts heuristic which essentially adds 2 for each pair of horizontally or vertically aligned times which need to pass each other.
-This heuristic is admissible and monotone.
+This heuristic is admissible, in fact it's monotone.
 Additional heuristics are generated randomly from $MD$, $LC$, and the misplaced tiles heuristic $MT$ which simply counts the number of non-blank tiles which are not already at their destinations.
@@
 We consider the case of multiple processes on a single multi-core machine, as well as multiple machines each running a single process.
-For comparison, we also include the performance of the single-threaded SMHA* algorithm that our work is based on.
+For comparison, we also include the performance of the sequential SMHA* algorithm that our work is based on.
 For that algorithm, we show performance as the number of heuristics it uses increases.
@@
   \caption{Running time}
-  \label{fig:singthread_runtime}
+  \label{fig1}
 \end{subfigure}
@@
   \caption{Solution length}
-  \label{fig:singthread_solnqual}
+  \label{fig2}
 \end{subfigure}
@@
   \caption{Running time}
-  \label{fig:numcores_runtime}
+  \label{fig3}
 \end{subfigure}
 \begin{subfigure}{3.2in}
-  \includegraphics[width=3.2in]{graphs/numcores_solutionquality}
+  \includegraphics[width=3.2in]{graphs/figure4}
   \caption{Solution length}
-  \label{fig:numcores_solnqual}
+  \label{fig4}
 \end{subfigure}
@@
     \hline
-    \multicolumn{4}{c}{Single-threaded SMHA* implementation} \\ \hline    
+    \multicolumn{4}{c}{Sequential SMHA* implementation} \\ \hline    
     \# cores & time & \# expand & soln len \\ \hline
@@
 
-% TODO: Should we have some kind of discussion on whether the parallelized version of SMHA* keeps the w1*w2 optimality guarantees that the single-threaded version does??
+% TODO: Should we have some kind of discussion on whether the parallelized version of SMHA* keeps the w1*w2 optimality guarantees that the sequential version does??
 Also, maybe part of this, but the longer we go without doing a table update sync-up, the more the algorithm resembles IMHA*.

### 2014-12-11 paper.tex
Experiments section.

--- 2014-12-11 paper.tex
@@
 
+Figure \ref{fig1} shows performance as we increase the number of processes (or heuristics, in the case of SMHA*).
+As shown, performance initially increases significantly, but then drops off as the number of processes increases.
+When running on a single machine, our distributed algorithm performs comparably to SMHA* for 4 or fewer cores, but decreases in performance as more processes are added.
+
+There are a number of likely explanations for this.
+First, the initial performance increase is largely a result of adding a heuristic which has more flexibility in returning a less-optimal path.
+Indeed, figure \ref{fig2} shows that the solution length increases considerably after adding just one non-anchor process.
+However, it is worth noting that our goal is specifically defined as trying to find a path within a specific multiple of the optimum.
+Thus, a longer path that remains within these bounds is viewed as an equally valid solution.
+The additional heuristics also provides more variety in the search, so that it is more likely that at least one of them succeeds in finding a good path.
+The drop in performance as more processes are added appears to be due to the increased communication cost of using a broadcast communication system.
+All processes must pause for this, and some might be slow in calling the broadcast command.
+Moreover, as more processes are added, the anchor receives more states that it must sift through, and the following message that it broadcasts increases in size as well.
+Thus, it appears that after a sufficiently diverse group of heuristics is added, the added communication cost of additional heuristics outweighs the benefits.
+These communication costs are analyzed in more detail later in this section and displayed in figure \ref{fig5}.
+
+Figure \ref{fig3} shows performance as we vary the number of iterations between communications.
+For reference, 100 iterations in our experiments is roughly 1-5 milliseconds.
+As one would expect, very frequent communication eventually causes the communication overhead to grow too large and hurts performance.
+On the other hand, as communication becomes extremely infrequent, the lack of shared information also hurts performance considerably.
+The optimal choice therefore depends heavily on the nature of the problem; in our case, it appears to be around 6,400 iterations.
+
+As discussed earlier, our algorithm does not appear to outperform SMHA* in the sliding tile puzzle domain.
+The most likely explanation for this is that because the heuristics we used for the puzzle can be calculated extremely efficiently, there are a huge number of expansions per unit of time.
+Each expanded state must be communicated to the anchor, which involves a considerable amount of work.
+It is likely that the amount of work in sending the states to the anchor, having the anchor process these states, broadcasting the best of those states to the other children, and having those children process the states, all together takes about as much work as expanding the state alone.
+Indeed, figure \ref{fig5} shows that as the number of processes increases, communication quickly becomes a huge fraction of the total time spent.
+This increased communication cost ultimately outweighs the benefits of additional processes.
+
+Nevertheless, figure \ref{fig4} displays what the results might look like in a problem where the heuristic takes longer to calculate.
+Specifically, we artificially slow the heuristic calculation by a factor of 10.
+The result is that fewer states are expanded per second, and thus the cost of communicating a state is much lower relative to the cost of expanding it.
+In this case, the results show that our distributed algorithm can leverage the benefit of additional cores to outperform SMHA* if we use the one-on-one communication method described earlier in this paper.
+
 \begin{figure}
@@
   \caption{Solution length}
-  \label{fig:singthread_solnqual}
+  \label{fig2}
 \end{subfigure}
-\caption{Comparison with Runtime performance and solution quality as the number of different heuristics (cores) on a single machine increases.}
+\caption{Comparison with Runtime performance and solution quality as the number of different heuristics (cores) on a single machine increases. Communication occurs every 5,000 iterations.}
 \label{fig:singthread}
@@
 \centering
-\begin{subfigure}{3.2in}
 \includegraphics[width=3.2in]{graphs/figure3}
-  \caption{Running time}
-  \label{fig:numcores_runtime}
-\end{subfigure}
-\begin{subfigure}{3.2in}
-  \includegraphics[width=3.2in]{graphs/numcores_solutionquality}
-  \caption{Solution length}
-  \label{fig:numcores_solnqual}
-\end{subfigure}
-\caption{Runtime performance and solution quality as the number of different heuristics (cores) on a single machine increases.}
-\label{fig:numcores}
+\caption{Runtime performance as communication frequency is varied. Experiment was done on 2 machines, with 2 processes on each machine.}
+\label{fig3}
 \end{figure}
@@
 \centering
-\begin{subfigure}{3.2in}
-  \includegraphics[width=3.2in]{graphs/nummachines_runningtime}
-  \caption{Running time}
-  \label{fig:nummachines_runtime}
-\end{subfigure}
-\begin{subfigure}{3.2in}
-  \includegraphics[width=3.2in]{graphs/nummachines_solutionquality}
-  \caption{Solution length}
-  \label{fig:nummachines_solnqual}
-\end{subfigure}
-\caption{Runtime performance and solution quality as the number of different machines (with one process each) increases from 1 to 16.}
-\label{fig:nummachines}
+\includegraphics[width=3.2in]{graphs/figure5}
+\caption{Cost of communication as the number of processes increases.
+Each process is on a different machine.
+Communication occurs every 5,000 iterations.
+Within MPI refers to only MPI commands.
+Outside MPI refers to other communication, such as serializing and reading states.}
+\label{fig5}
 \end{figure}
@@
 \centering
-\begin{subfigure}{2.0in}
-  \includegraphics[width=2.0in]{graphs/commfreq_runningtime}
-  \caption{Running time}
-  \label{fig:commfreq_runtime}
-\end{subfigure}
-\begin{subfigure}{2.0in}
-  \includegraphics[width=2.0in]{graphs/commfreq_expansions}
-  \caption{Number of state expansions}
-  \label{fig:commfreq_expansions}
-\end{subfigure}
-\begin{subfigure}{2.0in}
-  \includegraphics[width=2.0in]{graphs/commfreq_solutionquality}
-  \caption{Solution length}
-  \label{fig:nummachines_solnqual}
-\end{subfigure}
-\caption{Runtime performance, solution quality, and number of states expanded by the heuristic machines as the number of iterations between communication increases.
-*At the high end (12800 iterations between sync-ups), one of the four trials timed out (failed to find a solution within 5 minutes).
-The values in this graph are the average of the remaining three trials.}
-\label{fig:commfreq}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{ c | c | c | c }
-    \hline
-    \multicolumn{4}{c}{Single-threaded SMHA* implementation} \\ \hline    
-    \# cores & time & \# expand & soln len \\ \hline
-    1 & 1 & 2 & 3 \\ \hline
-    2 & 4 & 5 & 6 \\ \hline
-    4 & 7 & 8 & 9 \\ \hline
-    % FILL USING AMPERSAND-SEPARATED VALUES WITH "\\ \hline" APPENDED AT THE END
-
-     \multicolumn{4}{c}{Parallelized SMHA* (single machine, syncing every X iterations)} \\ \hline
-    \# cores & time & \# expand & soln len \\ \hline
-    % FILL USING AMPERSAND-SEPARATED VALUES WITH "\\ \hline" APPENDED AT THE END
-
-     \multicolumn{4}{c}{Distributed SMHA* (one process per machine, syncing every X iterations)} \\ \hline
-    \# machines & time & \# expand & soln len \\ \hline
-    % FILL USING AMPERSAND-SEPARATED VALUES WITH "\\ \hline" APPENDED AT THE END
-
-     \multicolumn{4}{c}{Distributed SMHA* with varying frequency of sync-ups} \\ \hline
-     \multicolumn{4}{c}{(4 processes across 2 machines)} \\ \hline
-    \# iterations & time & \# expand & soln len \\ \hline
-    % FILL USING AMPERSAND-SEPARATED VALUES WITH "\\ \hline" APPENDED AT THE END
-
-    \hline
-\end{tabular}
-\caption{Data from our trials}
-\label{fig:datatable}
+\includegraphics[width=3.2in]{graphs/figure4}
+\caption{Performance of distribution on heuristics that are 10x slower.}
+\label{fig4}
 \end{figure}

### 2014-12-11 paper.tex
changes

--- 2014-12-11 paper.tex
@@
 \cite{phillips2014pa} were able to achieve parallelization while retaining wA*'s guarantee of suboptimality without reexpansions.
-This method of parallelization makes inferences on when it is safe to expand multiple nodes in parallel, by resolving dependencies between potential optimal paths.
+This method of parallelization makes inferences on when it is safe to expand multiple states in parallel, by resolving dependencies between potential optimal paths.
 Their speed gains are dependent on the expansion operation being expensive enough to dominate the cost of computing these dependency checks, which may not be the case for all problems.
-By contrast, our method parallelizes the search process by simultaneously expanding nodes chosen by different heuristics, instead of multiple nodes chosen by a single heuristic.
+By contrast, our method parallelizes the search process by simultaneously expanding states chosen by different heuristics, instead of multiple states chosen by a single heuristic.
 We allow each state to be expanded by each heuristic at most once and, furthermore, no state will be expanded by a non-anchor heuristic which has received a message from another heuristic after the latter has expanded the same state.
@@
 Each machine runs its own heuristic, with the anchor machine's being the monotone heuristic.
-Each of the non-anchor heuristic machines periodically sends updates of the data structures (tables containing shortest path information found so far) to the anchor machine, which compares the data it receives and sends out table updates of the best paths known so far to all other nodes.
+Each of the non-anchor heuristic machines periodically sends updates of the data structures (tables containing shortest path information found so far) to the anchor machine, which compares the data it receives and sends out table updates of the best paths known so far to all other states.
 
@@
 The work was distributed over one or more of the 16-core machines in the Apt cluster.
+We associate one heuristic with each process that we run.
+In the following sections, we may use the terms machine and process interchangeably, even though we also tried running multiple heuristics (one per process) on one multi-core computer.
 % TODO: Machine specs here?
@@
 
-We chose MPI, despite is notorious complexity, because our algorithm requires extremely fast and efficient communication between the anchor and the child machines.
+We chose MPI, despite its complexity and low level of abstraction, because our algorithm requires extremely fast and efficient communication between the anchor and the child machines.
 Moreover, since the entire state must be replicated across all machines, the necessary communication was relatively straightforward.
@@
 Each child frequently checks for incoming messages using the \emph{MPI\_Iprobe} command; if a message is detected, the child receives the incoming data using \emph{MPI\_Recv} and responds with its own table of states.
-The anchor sends messages to each child one at a time, moving onto the next child when a response is received.
+The anchor sends messages to each child one at a time, moving on to the next child when a response is received.
 
@@
 After a pre-determined number of iterations, all child machines send their new states to the anchor using \emph{MPI\_Gather}.
-The anchor sifts through these responses, and keeps only the best g value for each state.
-Every state that has had a new best g value found is then broadcast to all the child nodes.
+The anchor sifts through these responses, and keeps only the best $g$-value for each state.
+Every state that has had a new best $g$-value found is then broadcast to all the child machines.
 
@@
 However, once called, the Broadcast command allows the anchor to only send the data to a few machines in the network.
-These machines, in turn, forward the data to other nodes in the machine, and so on.
-Thus, the anchor does not spend any more time communicating than any other node in the system, and the new state table is propagated in $O(\log(N))$ time.
-% TODO: Fill me in! MPI-specific feature...
+These machines, in turn, forward the data to other machines, and so on.
+Thus, the anchor does not spend any more time communicating than any other machine in the system, and the new state table is propagated in $O(\log(N))$ time.
 
-\subsubsection*{Update collation performed at anchor node}
+\subsubsection*{Update collation performed at anchor machine}
 
@@
 As sharing memory is infeasible in our distributed implementation, processes periodically broadcast their $g$ and $bp$ values to update one another.
-This can result in conflicting data for a particular state, which is always resolved by taking the data which gives the lowest $g$ value for that state as it contains the best estimate found so far.
+This can result in conflicting data for a particular state, which is resolved by taking the data which gives the lowest $g$ value for that state, as it contains the best estimate found so far.
 During conflict resolution, we also take the inclusive OR of the flags specifying whether the state has ever been expanded by the anchor, or by any machine at all.
+This prevents unnecessary reexpansions.
 
-Each broadcast by a heuristic machine only needs to contain the portion of the graph which has been updated by the machine since its most recent network communication.
+Each message only needs to contain the portion of the graph which has been updated by the sending machine since its most recent network communication.
 Nonetheless, if every machine broadcasts its updates directly to all other machines, the network would flood with messages and much conflict-resolution work would be duplicated.
-Therefore, we have every machine send its table updates exclusively to the anchor machine.
+Therefore, we instead have every machine send its table updates exclusively to the anchor machine.
 The anchor alone takes the burden of resolving all conflicts between duplicate states, and then periodically broadcasts the merged collection of updates to all machines.
@@
 To avoid unduly burdening the anchor, every machine is able to compute the anchor heuristic and includes its value in the state updates it sends to the anchor machine.
+Finally, the anchor machine's broadcasts also include its latest lower bound on the optimal solution cost, enabling child machines to deduce when they are below the $w_1w_2$ optimality factor.
 
@@
 $bp(s)$ represents the parent of $s$ along the optimal path found so far from the start to $s$.
-As a single pointer cannot be shared across a distributed replicated data structure, our first implementation used a full state representation (5x5 sliding tile puzzle) in place of $bp(s)$, nearly doubling the length of network messages.
-However, notice that $bp(s)$ is not needed to perform the search and compute $g$-values; it is only needed to reconstruct the solution path once the search terminates.
+As a single pointer cannot be shared across a distributed replicated data structure, our first implementation used a full state representation (5x5 puzzle configuration) in place of $bp(s)$, nearly doubling the length of network messages.
+However, notice that $bp(s)$ is not needed to perform the search $g$-value computations; it is only needed to reconstruct the solution path once the search terminates.
 
@@
 At each step, we move from the current state $s$ to the neighbor $s'$ which minimizes $g(s') + c(s',s)$ among the possible parents $s'=bp(s)$ pooled from all machines.
-Although we still end up sending full state representations, these are only for the parents of states along the path.
+Although we still end up sending full state representations, these are only for the parents of states along the path, rendering the overhead negligible.
 

### 2014-12-11 paper.tex
adding the hamstar since it's never mentioned in the report
Conclusion + Abstract

@@
 For this project we implemented a distributed version of MHA*, with one process per heuristic.
-We show...
-[performance]}
-%TODO: Highlight important results
+We compare the results of our distributed implementation to the original MHA* algorithm.
+In particular, we provide evidence that in domains where state expansion is expensive, such as when heuristic calculations are slow, our distributed approach may outperform MHA*.
 
@@
 In the following sections, we may use the terms machine and process interchangeably, even though we also tried running multiple heuristics (one per process) on one multi-core computer.
-% TODO: Machine specs here?
-Or in the experimental setup section?
 
@@
 These heuristics are not at all admissible, and their high values give the search a goal-directed focus while the anchor is underestimating the true distance by a large margin.
-
-% TODO: Apt? Machine specs?
 
@@
 
+In this paper we outlined the details of a distributed version of the SMHA* algorithm.
+We detailed the structure of the algorithm, and presented experimental results comparing the performance of the algorithm while varying a number of parameters.
+We also compared its performance to the original SMHA* algorithm.
+Although we found that in the specific domain we tested on (the sliding tile puzzle) our distributed variant did not outperform the original algorithm, we provided evidence that our algorithm could outperform SMHA* in domains where heuristics are difficult to calculate, or in general where state expansion is expensive.
 % TODO: Fill me in!

### 2014-12-11 paper.tex
Results solidification

--- 2014-12-11 paper.tex
@@
 Figure \ref{fig1} shows performance as we increase the number of processes (or heuristics, in the case of SMHA*).
-As shown, performance initially increases significantly, but then drops off as the number of processes increases.
+As shown, performance increases significantly at first, but then drops off as the number of processes increases.
 When running on a single machine, our distributed algorithm performs comparably to SMHA* for 4 or fewer cores, but decreases in performance as more processes are added.
@@
 There are a number of likely explanations for this.
-First, the initial performance increase is largely a result of adding a heuristic which has more flexibility in returning a less-optimal path.
+First, the initial performance increase is largely a result of adding a heuristic with a greedier bias, typically finding a slightly less optimal path more quickly.
 Indeed, figure \ref{fig2} shows that the solution length increases considerably after adding just one non-anchor process.
-However, it is worth noting that our goal is specifically defined as trying to find a path within a specific multiple of the optimum.
-Thus, a longer path that remains within these bounds is viewed as an equally valid solution.
-The additional heuristics also provides more variety in the search, so that it is more likely that at least one of them succeeds in finding a good path.
+However, it is worth noting that our criterion is specifically to find a path whose cost is within the specified multiple (in our case, 4) of the optimum.
+Crucially, the combination of monotone and inadmissible search is able to guarantee the near-optimality of the solution; thus, the slightly longer path is viewed as an equally valid solution.
+The additional heuristics provide more variety in the search, making it more likely for at least one of them (or some combination) to succeed in finding a good path.
 The drop in performance as more processes are added appears to be due to the increased communication cost of using a broadcast communication system.
@@
 Thus, it appears that after a sufficiently diverse group of heuristics is added, the added communication cost of additional heuristics outweighs the benefits.
-These communication costs are analyzed in more detail later in this section and displayed in figure \ref{fig5}.
+These communication costs are analyzed in more detail later in this section and are displayed in figure \ref{fig5}.
 
 Figure \ref{fig3} shows performance as we vary the number of iterations between communications.
-For reference, 100 iterations in our experiments is roughly 1-5 milliseconds.
+For reference, 100 iterations in our experiments takes roughly 1-5 milliseconds.
 As one would expect, very frequent communication eventually causes the communication overhead to grow too large and hurts performance.
 On the other hand, as communication becomes extremely infrequent, the lack of shared information also hurts performance considerably.
-The optimal choice therefore depends heavily on the nature of the problem; in our case, it appears to be around 6,400 iterations.
+The optimal communication interval therefore depends heavily on the nature of the problem; in our case, it appears to be around 6,400 iterations.
 
 As discussed earlier, our algorithm does not appear to outperform SMHA* in the sliding tile puzzle domain.
-The most likely explanation for this is that because the heuristics we used for the puzzle can be calculated extremely efficiently, there are a huge number of expansions per unit of time.
+The most likely explanation for this is that, because the successor states and heuristics we used for the puzzle can be calculated extremely efficiently, there are a huge number of expansions per unit time.
 Each expanded state must be communicated to the anchor, which involves a considerable amount of work.
-It is likely that the amount of work in sending the states to the anchor, having the anchor process these states, broadcasting the best of those states to the other children, and having those children process the states, all together takes about as much work as expanding the state alone.
+It is likely that the amount of work in sending the states to the anchor, having the anchor process these states, broadcasting the best of those states to the other children, and having those children process the states, altogether takes about as much work as expanding the state alone.
 Indeed, figure \ref{fig5} shows that as the number of processes increases, communication quickly becomes a huge fraction of the total time spent.
@@
 The result is that fewer states are expanded per second, and thus the cost of communicating a state is much lower relative to the cost of expanding it.
-In this case, the results show that our distributed algorithm can leverage the benefit of additional cores to outperform SMHA* if we use the one-on-one communication method described earlier in this paper.
+In this case, our distributed algorithm can leverage the benefit of additional cores to outperform SMHA* if we use the one-on-one communication method described earlier in this paper.
+Thus, a distributed approach may prove helpful in domains such as manipulator arm planning, in which expansions are slowed by the collision checking and numerical integration needed to compute successor states and heuristics.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=1.6in]{graphs/hamstar}
+\end{figure}
+
 \end{document}

### 2014-12-11 paper.tex
Expanded conclusion

--- 2014-12-11 paper.tex
@@
 Thus, the anchor does not spend any more time communicating than any other machine in the system, and the new state table is propagated in $O(\log(N))$ time.
+
+We ultimately chose the broadcast model for our implementation due to the more balanced workload on the anchor.
+In fact, in the specific problem we analyzed (the sliding tile puzzle), using the one-on-one communication model was not a viable option; it would take longer for the anchor to process incoming states and send them out than it would for child nodes to expand new states.
 
@@
 We also compared its performance to the original SMHA* algorithm.
+
+In the specific domain we tested on (the sliding tile puzzle) our distributed variant did not outperform the original algorithm.
+The culprit appears to be the excessive communication cost relative to the cost of expanding a state.
+We were partially able to mitigate this by using the broadcast model of communication, and by reducing the footprint of a communicated state by not recovering the backpointer until the end and by offloading the calculation of the anchor heuristic onto the child machines.
+Nevertheless, it appears more optimizations would be required in order to gain a significant benefit over vanilla SMHA*.
+One possibility is the use of asynchronous communication.
+Both of our communication models required the sender and receiver to be ready to communicate simultaneously.
+If one was ready before the other, then that machine would be blocked until the counterpart responded.
+Using asynchronous communication would eliminate the requirement that all machines cease work until all are ready, which becomes a more significant problem as the size of the distribution grows due to the possibility of slow outlier machines.
+This would also make the algorithm more robust to failure, since the algorithm could proceed even if one machine becomes unresponsive.
+The latest version of MPI (MPI 3.0) appears to include an asynchronous broadcast option that would accomplish something similar to this.
+
 Although we found that in the specific domain we tested on (the sliding tile puzzle) our distributed variant did not outperform the original algorithm, we provided evidence that our algorithm could outperform SMHA* in domains where heuristics are difficult to calculate, or in general where state expansion is expensive.
-% TODO: Fill me in!
 

### 2014-12-11 paper.tex
results stuff

--- 2014-12-11 paper.tex
@@
 \begin{figure}
-\centering \includegraphics[width=3in]{pseudocode}
-\caption{Pseudocode from Aine et al., detailing the original sequential version of the algorithm we parallelized.}
+\centering
+\begin{subfigure}{3in}
+  \includegraphics[width=3in]{pseudocode}
+  \caption{Pseudocode}
+  \label{fig1}
+\end{subfigure}
+\begin{subfigure}{2in}
+  \includegraphics[width=2in]{graphs/puzzle}
+  \caption{15-puzzle goal configuration}
+  \label{fig2}
+\end{subfigure}
+\caption{Pseudocode from Aine et al., detailing the original sequential version of the algorithm we parallelized, alongside the sliding tile puzzle which we use it to solve.}
 \label{fig:pseudocode}
@@
 The result is that fewer states are expanded per second, and thus the cost of communicating a state is much lower relative to the cost of expanding it.
-In this case, our distributed algorithm can leverage the benefit of additional cores to outperform SMHA* if we use the one-on-one communication method described earlier in this paper.
+In this case, our distributed algorithm can leverage the benefit of additional cores to outperform SMHA* if we use the one-on-one communication method described earlier in this report.
 Thus, a distributed approach may prove helpful in domains such as manipulator arm planning, in which expansions are slowed by the collision checking and numerical integration needed to compute successor states and heuristics.
@@
 
-In this paper we outlined the details of a distributed version of the SMHA* algorithm.
-We detailed the structure of the algorithm, and presented experimental results comparing the performance of the algorithm while varying a number of parameters.
-We also compared its performance to the original SMHA* algorithm.
-Although we found that in the specific domain we tested on (the sliding tile puzzle) our distributed variant did not outperform the original algorithm, we provided evidence that our algorithm could outperform SMHA* in domains where heuristics are difficult to calculate, or in general where state expansion is expensive.
+In this report, we outlined the details of a distributed version of the SMHA* algorithm.
+We detailed the implementation of the algorithm, and presented experimental results comparing its performance while varying a number of parameters.
+We also compared its performance to the original SMHA*.
+Although we found that our distributed variant did not outperform the original SHMA* in the specific domain we tested on (the sliding tile puzzle), we provided evidence that our algorithm could outperform SMHA* in domains where heuristics are difficult to calculate, or in general where state expansion is expensive.
 % TODO: Fill me in!
@@
 
-%\section{Final Paper Outline}
+%\section{Final Report Outline}
 
@@
 
-\begin{figure}
+\begin{figure}[b]
 \centering

### 2014-12-11 paper.tex
Update

--- 2014-12-11 paper.tex
@@
 
-In this paper we outlined the details of a distributed version of the SMHA* algorithm.
-We detailed the structure of the algorithm, and presented experimental results comparing the performance of the algorithm while varying a number of parameters.
+In this report, we outlined the details of a distributed version of the SMHA* algorithm.
+We detailed the implementation of the algorithm, and presented experimental results comparing its performance while varying a number of parameter.
 We also compared its performance to the original SMHA* algorithm.
 
-In the specific domain we tested on (the sliding tile puzzle) our distributed variant did not outperform the original algorithm.
+In the specific domain we tested on (the sliding tile puzzle), our distributed algorithm did not outperform the original SHMA*.
 The culprit appears to be the excessive communication cost relative to the cost of expanding a state.

### 2014-12-11 paper.tex
abstract and conclusion

@@
 \abstract{A* is a popular graph search algorithm used in many artificial intelligence applications.
-The original A* algorithm makes use of a single heuristic to prioritize exploration through a graph, but a variation on the algorithm known as multi-heuristic A* (MHA*) makes use of multiple heuristics to more quickly arrive at approximate solutions.
-For this project we implemented a distributed version of MHA*, with one process per heuristic.
-We compare the results of our distributed implementation to the original MHA* algorithm.
-In particular, we provide evidence that in domains where state expansion is expensive, such as when heuristic calculations are slow, our distributed approach may outperform MHA*.
+The original A* algorithm uses a single heuristic to prioritize exploration through a graph, but a variation on the algorithm known as shared multi-heuristic A* (SMHA*) makes use of multiple heuristics to more quickly arrive at near-optimal solutions.
+For this project, we implemented a distributed version of SMHA*, separating each heuristic into a separate process.
+We compare the results of our distributed implementation to the original SMHA* algorithm.
+In particular, we provide evidence that in domains where state expansions are expensive, our distributed approach may outperform SMHA*.
 
@@
 In this report, we outlined the details of a distributed version of the SMHA* algorithm.
-We detailed the implementation of the algorithm, and presented experimental results comparing its performance while varying a number of parameter.
+We detailed the implementation of the algorithm, and presented experimental results comparing its performance while varying a number of parameters.
 We also compared its performance to the original SMHA* algorithm.
@@
 If one was ready before the other, then that machine would be blocked until the counterpart responded.
-Using asynchronous communication would eliminate the requirement that all machines cease work until all are ready, which becomes a more significant problem as the size of the distribution grows due to the possibility of slow outlier machines.
-This would also make the algorithm more robust to failure, since the algorithm could proceed even if one machine becomes unresponsive.
+Using asynchronous communication would eliminate the requirement that all machines cease work until all are ready which, due to the likelihood of slow outlier machines, becomes a more significant problem as the machine count increases.
+This would also make the algorithm more robust to failure, since the algorithm could proceed even if some machines become unresponsive.
 The latest version of MPI (MPI 3.0) appears to include an asynchronous broadcast option that would accomplish something similar to this.
 
-Although we found that in the specific domain we tested on (the sliding tile puzzle) our distributed variant did not outperform the original algorithm, we provided evidence that our algorithm could outperform SMHA* in domains where heuristics are difficult to calculate, or in general where state expansion is expensive.
+Although our distributed algorithm did not outperform SHMA* on the sliding tile puzzle, we have provided evidence that our algorithm could outperform SMHA* in domains where heuristics are difficult to calculate, or in general where state expansion is expensive.
 
@@
 \centering
-\includegraphics[width=1.6in]{graphs/hamstar}
+\includegraphics[width=0.42in]{graphs/hamstar}
 \end{figure}

### 2014-12-11 paper.tex
Updated paper stuff

--- 2014-12-11 paper.tex
@@
 We investigated other frameworks designed for iterative algorithms on a stable dataset such as Spark and GraphLab.
-However, most of these frameworks were designed for chunking up a large dataset upon which a single operation is performed.
+However, most of these frameworks were designed for chunking up a large dataset and assigning pieces to workers.
 Our application's needs differed from this model in that the dataset being operated on (the graph being explored) was shared by all workers, and the workers themselves were performing differing operations on subsets of the same data.
@@
 Moreover, since the entire state must be replicated across all machines, the necessary communication was relatively straightforward.
-Finally, we found detailed documentation on MPI, explaining the low-level workings of each command, much more readily accessible than comparable documentation on Spark or GraphLab, which seemed more focused on abstracting the details of the system away from the programmer.
+Finally, we found that detailed documentation on MPI, explaining the low-level workings of each command, was much more readily accessible than comparable documentation on Spark or GraphLab, which seemed more focused on abstracting the details of the system away from the programmer.
 Given our emphasis on speed, MPI seemed like an appropriate choice.
@@
 We considered two communication models for our algorithm.
-The first one has the anchor periodically send each child a table of states.
+The first one, which we call the one-on-one model, has the anchor periodically send all children a table of states.
 The table contains information for new states that were either received by the anchor, or calculated by the anchor, since the last time the anchor communicated with this child.
@@
 The anchor sends messages to each child one at a time, moving on to the next child when a response is received.
+Once the data has been sent to all children, the anchor will run its own iterations until the next cycle should begin.
 
-The second communication model involves a period broadcast using the \emph{MPI\_Bcast} command.
+The second communication model involves a periodic broadcast using the \emph{MPI\_Bcast} command.
 After a pre-determined number of iterations, all child machines send their new states to the anchor using \emph{MPI\_Gather}.
@@
 
-The one-on-one communication model has the benefit that it does not require all machines to cease work while communication happens; only the anchor and the single child machine need to participate in the communication.
+The one-on-one communication model has the benefit that it does not require all machines to cease work while communication happens; only the anchor and a single child machine need to participate in the communication.
 The downside is that the anchor must then do $O(N)$ communications, where $N$ is the number of children, and the time it takes to propogate the new state tables across the entire system is also $O(N)$.

### 2014-12-11 paper.tex
broadcast section

@@
 We considered two communication models for our algorithm.
-The first one, which we call the one-on-one model, has the anchor periodically send all children a table of states.
-The table contains information for new states that were either received by the anchor, or calculated by the anchor, since the last time the anchor communicated with this child.
-Each child frequently checks for incoming messages using the \emph{MPI\_Iprobe} command; if a message is detected, the child receives the incoming data using \emph{MPI\_Recv} and responds with its own table of states.
+The first, which we call the one-on-one model, has the anchor periodically send all children a table of states.
+The table contains information for new states that were either received or calculated by the anchor since its most recent communication with this child.
+Each child frequently checks for incoming messages using the \emph{MPI\_Iprobe} command: if a message is detected, the child receives the incoming data using \emph{MPI\_Recv} and responds with its own table of states.
 The anchor sends messages to each child one at a time, moving on to the next child when a response is received.
-Once the data has been sent to all children, the anchor will run its own iterations until the next cycle should begin.
+Once the data has been sent to all children, the anchor will run its own expansion iterations until a time when the next cycle should begin.
 
 The second communication model involves a periodic broadcast using the \emph{MPI\_Bcast} command.
-After a pre-determined number of iterations, all child machines send their new states to the anchor using \emph{MPI\_Gather}.
+After a pre-determined number of iterations, all child machines send their new or updated states to the anchor using \emph{MPI\_Gather}.
 The anchor sifts through these responses, and keeps only the best $g$-value for each state.
@@
 These machines, in turn, forward the data to other machines, and so on.
-Thus, the anchor does not spend any more time communicating than any other machine in the system, and the new state table is propagated in $O(\log(N))$ time.
+Thus, the anchor does not spend significantly more time communicating than other machines in the system, and the new state table is propagated in $O(\log(N))$ time.
 
 We ultimately chose the broadcast model for our implementation due to the more balanced workload on the anchor.
-In fact, in the specific problem we analyzed (the sliding tile puzzle), using the one-on-one communication model was not a viable option; it would take longer for the anchor to process incoming states and send them out than it would for child nodes to expand new states.
+In fact, for the case of the sliding tile puzzle, the one-on-one communication model was not a viable option; it would take longer for the anchor to process incoming states and send them out than it would for child machines to expand new states.
 

### 2014-12-11 paper.tex
Final changes
Changes?
boom

@@
 We tested our algorithm on the 24-puzzle, a 5x5 grid of sliding blocks which must be rearranged to achieve a standard goal configuration, somewhat akin to solving a Rubix cube.
-Uniformly at random, we generated ten starting configurations among the over $3\times 10^{23}$ states of the puzzle.
+Uniformly at random, we generated 10 starting configurations among the over $3\times 10^{23}$ states of the puzzle.
 The data we collected were (1) running time, (2) total number of states in the problem graph that were expanded across all processes during the search, and (3) the length in steps of the best solution at termination.
+For each of these, we discard the top and bottom value.
+Our plots show the min, mean and max of the 8 remaining values.
 Our main evaluation criterion is the time to complete the search with the required suboptimality guarantee, which we fix to $w_1 w_2 = 4$.
@@
 We repeat this process as many times as we require non-anchor heuristics.
-These heuristics are not at all admissible, and their high values give the search a goal-directed focus while the anchor is underestimating the true distance by a large margin.
-
-% TODO: Talk about what we varied for the tests.
-Frequency of communication, effect of adding more heuristics, effect of adding more machines (maybe we want a graph that compares the effect of adding more heuristics on the same machine vs.
-adding more heuristics by using more machines), and comparison to single thread/round robing running time
+These heuristics are not at all admissible, and their high values give the search a goal-directed focus as the anchor typically underestimates the true distance by a large margin.
 
@@
 Therefore, SMHA* guarantees that the solution found at termination is at most $w_1 w_2$ times the optimum value.
+This guarantee applies equally well in our distributed variant.
 For all experiments, we used $w_1 = w_2 = 2$.
@@
 
-Figure \ref{fig1} shows performance as we increase the number of processes (or heuristics, in the case of SMHA*).
+Figure \ref{fig1} plots runtime as we increase the number of processes (or heuristics, in the case of SMHA*).
 As shown, performance increases significantly at first, but then drops off as the number of processes increases.
@@
 \includegraphics[width=3.2in]{graphs/figure4}
-\caption{Performance of distribution on heuristics that are 10x slower.}
+\caption{Performance of distribution on heuristics that are 10x slower. Since each iteration is slower, we communicate every 1,000 iterations instead of every 5,000 for this experiment.}
 \label{fig4}
@@
 
-% TODO: Should we have some kind of discussion on whether the parallelized version of SMHA* keeps the w1*w2 optimality guarantees that the sequential version does??
-Also, maybe part of this, but the longer we go without doing a table update sync-up, the more the algorithm resembles IMHA*.
-It might be good to point this out somewhere, maybe in the Results/Discussion section
-
-%\section{Final Report Outline}
-
-%\subsection{Introduction}
-%\subsection{Related Work}
-%\subsection{System Description}
-%\subsection{Evaluation}
-%\subsection{Conclusions and Future Work}
-
 \bibliographystyle{ieeetr}
@@
 \centering
-\includegraphics[width=0.42in]{graphs/hamstar}
+\includegraphics[width=0.4in]{graphs/hamstar}
 \end{figure}

### 2014-12-15 paper.tex
abstract is no longer the same size as everything else
some cleanup, but for some reason the prints in graphfile2.m stopped working

@@
-\documentclass[11pt]{article}
+\documentclass{article}
 
@@
 
-\abstract{A* is a popular graph search algorithm used in many artificial intelligence applications.
+\begin{abstract}
+A* is a popular graph search algorithm used in many artificial intelligence applications.
 The original A* algorithm uses a single heuristic to prioritize exploration through a graph, but a variation on the algorithm known as shared multi-heuristic A* (SMHA*) makes use of multiple heuristics to more quickly arrive at near-optimal solutions.
@@
 In particular, we provide evidence that in domains where state expansions are expensive, our distributed approach may outperform SMHA*.
+\end{abstract}
 
@@
 
-\begin{figure}
-\centering \includegraphics[width=5.0in]{system-diagram}
+\begin{figure}[h]
+\centering \includegraphics[width=5.0in]{graphs/system-diagram}
 \caption{A diagram of our system. One machine starts up the others and runs the monotone anchor search. The other machines run their individual heuristics. Communication is done using MPI. 
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
 \begin{subfigure}{3in}
-  \includegraphics[width=3in]{pseudocode}
+  \includegraphics[width=3in]{graphs/pseudocode}
   \caption{Pseudocode}
@@
 The first weight ($w_1$) gives the search a more depth-first goal-directed focus.
-The second weight ($w_2$) determines the margin by which a proposed solution's cost may exceed the anchor's minimum frontier key value, which is itself $w1$ times a lower bound on the optimum solution cost.
+The second weight ($w_2$) determines the margin by which a proposed solution's cost may exceed the anchor's minimum frontier key value, which is itself $w_1$ times a lower bound on the optimum solution cost.
 Therefore, SMHA* guarantees that the solution found at termination is at most $w_1 w_2$ times the optimum value.
@@
 \centering
-\includegraphics[width=0.4in]{graphs/hamstar}
+\includegraphics[width=0.49in]{graphs/hamstar}
 \end{figure}
